### 2023-01-18 manuscript.Rmd
initial draft of some ideas

---
title             : "A commentary to 'Responsible Research Assessment II: A specific proposal for hiring and promotion in psychology'"
shorttitle        : "Responsible research assessment"

author: 
  - name          : "Andreas M. Brandmaier"
    affiliation   : "1,2,3"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
      - "Supervision"
  - name          : "Maximilian Ernst"
    affiliation   : "2"
    role:
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Aaron Peikert"
    affiliation   : "2"
    role:
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "MSB Medical School Berlin"
  - id            : "2"
    institution   : "Max Planck Institute for Human Development"
  - id            : "3"
    institution   : "Humboldt Universität zu Berlin"

abstract: |
  TBD

keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes (Schönbrodt et al., X), Gaertner et al. suggest a specific implementation to broaden the range of what constitutes research contributions. Gaertner et al. provide evaluation schemes for published manuscripts, published datasets, and software contributions. Here, we specifically comment on the proposed evaluation scheme for research software contributions (Table 3). In the following, we comment on dimensions metrics, lines of code, software engineering, impact, peer-review, usability.

## Metrics

Gaertner et al. suggest the number of citations (their indicator #9 in Table 3) to evaluate the impact of a software package. They acknowledge the problem that citations inconsistently point to software packages and/or accompanying software packages. For example, one of the authors maintains the pdc package for R. Google scholar counts X citations to the package and X citations to the accompanying paper, of which XX citations cite both, which permits a simple addition of the citation counts.

Date of first fully functional public version is a central field in the proposal as it serves as a denominator in computing citations relative to software age. In our own experience, research software often evolves very differently to the standards that software engineering prescribes. In best-practice software development, the first step is software specification or requirements engineering. This is the process of understanding and defining what features and services are required from the product and how they are to be offered. This involves research on existing solutions, discussions with users, user interface testing with mockups or user observations while testing prototypes. A requirement specification documents the expected functionality, such that development time and cost can be estimated and requirements can be formally checked with the final product for consistency and completeness. From our own experience, hardly any of these steps apply to scientific software development. Scientific software is often developed as a side-project fuelled by enthusiasm instead of proper funding. Some pieces of software start as testbed implementation of a newly proposed methodology, others come to life as wrappers or graphical interfaces for other packages, some others were never meant to be shared but are so useful that they became a package on their own.

We strongly object to the metric Github stars. First of all, it is tied only to a single, commercial platform and allows no freedom to choose free alternatives such as local Gitlab instances or other platforms. More importantly, we question the validity of such metrics that like rating systems on commercial platforms are too easy to be gamed by creating fake accounts that provide high ratings.

In our opinion, total citation metrics, number of users, downloads per month and similar may provide a coarse measure for the impact of a software package, even though it is important to note that these metrics are difficult to compare across different types of software package. However, we would like to point out that the request for citation metrics with respect to software packages seems inconsistent with the authors condemnation of citation metrics for publications.

Complementing the quantitative aspect of impact, we suggest to introduce a qualitative response. Researchers should indicate at least one (and up to three) empirical research projects that directly benefitted from the software contributions they made. We believe this is a fair assessment of the real impact no matter whether someone contributed documentation/tutorial to run certain types of models, whether someone contributed a specific feature that enabled a given research project to run their analysis, or whether someone is the author of the entire package.

## Software Engineering standards

As mentioned before, software engineering standards are often not adhered to in the ad-hoc
development of scientific software. Thus, at the minimum, we would like to encourage researchers if they provide unit tests. Unit tests are the backbone of software quality and testing. Developing unit tests may often cost more time than developing the actual code. At the same time, they are usually unrecognized because they are invisible to the end user and provide no functionality on its own.

## Lines of code

We strongly oppose the lines of code (LOC) metric that was proposed to measure effort. Lines of code is a common metric in software development useful to gauge the size of an entire code base or its change over time. There are several problems in using this indicator for effort. First of all, lines of code is an ambiguous metric. It can be calculated differently depending on whether comments do count or whether physical lines of code or logical lines of code are counted.  Second, lines of codes for code contributions highly depends on the programming language, mastery and personal programming style. In particular, large LOC of a researcher may simply mean that they write unefficient and repetitive code, one of the great sins of programming. A feature of good software is modularity because it enable reuse of functions both inside and outside the project. A great example is the CRAN library that enables packages to built upon other packages and reuse their functionality. However, if LOC were a metric that a researcher are compared with, it encourages researchers to copy&paste chunks of code from other packages instead of importing them in their project.

## Peer-Review

Similar to peer-review for manuscripts, software code is amenable to peer-review. It is common
practice in corporate software development, to have peers review code. For example, if a code base is changed, it is good practice to request a change with the developer team before actually changing the code base. The change is reviewed by experts and/or maintainers before it is integrated in the active branch of development. In developing these guidelines, we see the opportunity to foster such good practices and reward researchers that actively engage in reviewing computer code. Software peer-review is increasingly becoming recognized as a means to enable better research software. Peer-review can happen in the context of dedicated software journal (e.g, Journal of Statistical Software), as part of the standard review process (for guidelines, see https://www.carlboettiger.info/2013/06/13/what-I-look-for-in-software-papers.html), or as part of submissions to software archives that guarantuee certain standards (e.g., rOpenSci). We suggest that active participation in the review of scientific software, independent of the involvement or maintance of specific software products, should be a dimension of positive evaluation.

## Reproducibility

->Aaron and Max

## Summary

In sum, we are thankful to Gaertner et al. for highlighting the importance of research software contributions as academically valuable products. Specifically, we value ... However, we are critical of ... for the reasons mentioned here. We propose to ...

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

### 2023-01-18 manuscript.Rmd
some refs added

@@
 
-Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes (Schönbrodt et al., X), Gaertner et al.
-suggest a specific implementation to broaden the range of what constitutes research contributions.
+Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes (Schönbrodt et al., X), @gartner2022responsible suggest a specific implementation to broaden the range of what constitutes research contributions.
 Gaertner et al.
@@
 Here, we specifically comment on the proposed evaluation scheme for research software contributions (Table 3).
-In the following, we comment on dimensions metrics, lines of code, software engineering, impact, peer-review, usability.
+In the following, we comment on the dimensions metrics, lines of code, software engineering, impact, peer-review, usability.
 
@@
 
-Gaertner et al.
-suggest the number of citations (their indicator #9 in Table 3) to evaluate the impact of a software package.
+@gartner2022responsible suggest the number of citations (their indicator #9 in Table 3) to evaluate the impact of a software package.
 They acknowledge the problem that citations inconsistently point to software packages and/or accompanying software packages.
@@
 
-In our opinion, total citation metrics, number of users, downloads per month and similar may provide a coarse measure for the impact of a software package, even though it is important to note that these metrics are difficult to compare across different types of software package.
-However, we would like to point out that the request for citation metrics with respect to software packages seems inconsistent with the authors condemnation of citation metrics for publications.
+In our opinion, total citation metrics, number of users, downloads per month and similar may provide a coarse measure for the impact of a software package, even though it is important to note that these metrics are difficult to compare across different types of software packages.
+For example, it is probably more commonly accepted to cite data analysis packages (particularly if they are commercial packages) than supporting packages (such as papaja [@R-papaja] used to render this article).
+A further complication arises through the fact that functionality of successful, modular scientific software is imported, re-used, and re-exported in other software packages.
+While we highly encourage this from both a software engineering perspective and for scientific progress, it poses further difficulties to citation-based metrics.
+For example, the pdc package [@pdc-paper] offers functions to cluster time-series based on relative complexity.
+The TSclust package [@TSclust] is a wrapper package that unites several divergence measures for time series clustering conveniently in a single package, particularly, the pdc clustering approach.
+This potentially leads to researchers only citing the TSclust package but not the original algorithm (e.g., as in the work by @staroverova).
 
-Complementing the quantitative aspect of impact, we suggest to introduce a qualitative response.
-Researchers should indicate at least one (and up to three) empirical research projects that directly benefitted from the software contributions they made.
+Last, we would like to point out that the request for citation metrics with respect to software packages seems inconsistent with the authors condemnation of citation metrics for publications.
+
+Complementing the quantitative aspect of impact, we suggest to introduce a mandatory qualitative judgement on the impact of the software, which fits with what the others proposed as indicator "merit / impact statement".
+In our view, researchers should be requested to indicate at least one (and up to three) empirical research projects that directly benefitted from the software contributions they made.
 We believe this is a fair assessment of the real impact no matter whether someone contributed documentation/tutorial to run certain types of models, whether someone contributed a specific feature that enabled a given research project to run their analysis, or whether someone is the author of the entire package.

### 2023-01-18 manuscript.Rmd
fixed refs

@@
 
-Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes (Schönbrodt et al., X), @gartner2022responsible suggest a specific implementation to broaden the range of what constitutes research contributions.
+Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggest a specific implementation to broaden the range of what constitutes research contributions.
 Gaertner et al.

### 2023-01-19 manuscript.Rmd
added to intro on research software engineering and importance of software in science

@@
 provide evaluation schemes for published manuscripts, published datasets, and software contributions.
-Here, we specifically comment on the proposed evaluation scheme for research software contributions (Table 3).
-In the following, we comment on the dimensions metrics, lines of code, software engineering, impact, peer-review, usability.
+Why are research software contributions important academic contributions? We would like to respond with a quote from the Science Code Manifesto (http://sciencecodemanifesto.org/):
+"Software is an essential research product, and the effort to produce, maintain, adapt, and 
+curate code must be recognized."
+and
+"Software is a cornerstone of science. Without software, twenty-first century science would be impossible."
+Yet, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivice good software engineering practice that we would expect from a research software engineer [@baxter2012research].
+ Here, we specifically comment on the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible).
+We comment on the dimensions metrics, lines of code, software engineering, impact, peer-review, usability.
 

### 2023-01-23 manuscript.Rmd
fixed some language issues

@@
 
-Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggest a specific implementation to broaden the range of what constitutes research contributions.
+Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a specific implementation of these principles to broaden the range of what constitutes research contributions.
 Gaertner et al.
 provide evaluation schemes for published manuscripts, published datasets, and software contributions.
-Why are research software contributions important academic contributions? We would like to respond with a quote from the Science Code Manifesto (http://sciencecodemanifesto.org/):
+We specifically endorse the increased emphasis on research software as recognizable and laudable scientific contribution.
+Why are research software contributions important scientific contributions? We would like to respond with a quote from the Science Code Manifesto (http://sciencecodemanifesto.org/):
 "Software is an essential research product, and the effort to produce, maintain, adapt, and 
@@
 "Software is a cornerstone of science. Without software, twenty-first century science would be impossible."
-Yet, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivice good software engineering practice that we would expect from a research software engineer [@baxter2012research].
+Yet, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivise software development and, specifically, good software engineering practice that we would expect from a research software engineer [@baxter2012research].
  Here, we specifically comment on the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible).
@@
 
+<!-- This needs to be shortened a lot-->
 Date of first fully functional public version is a central field in the proposal as it serves as a denominator in computing citations relative to software age.
@@
 
-Complementing the quantitative aspect of impact, we suggest to introduce a mandatory qualitative judgement on the impact of the software, which fits with what the others proposed as indicator "merit / impact statement".
-In our view, researchers should be requested to indicate at least one (and up to three) empirical research projects that directly benefitted from the software contributions they made.
+Complementing the quantitative aspect of impact, we suggest to introduce a mandatory qualitative judgement on the impact of the software, which fits with what @gartner2022responsible proposed as indicator "merit / impact statement".
+In our view, researchers should be requested to indicate at least one (and up to three) empirical research projects that directly benefited from the software contributions they made.
 We believe this is a fair assessment of the real impact no matter whether someone contributed documentation/tutorial to run certain types of models, whether someone contributed a specific feature that enabled a given research project to run their analysis, or whether someone is the author of the entire package.
@@
 
-We strongly oppose the lines of code (LOC) metric that was proposed to measure effort.
+We discourage the lines of code (LOC) metric that was proposed to measure effort.
 Lines of code is a common metric in software development useful to gauge the size of an entire code base or its change over time.
@@
 Second, lines of codes for code contributions highly depends on the programming language, mastery and personal programming style.
-In particular, large LOC of a researcher may simply mean that they write unefficient and repetitive code, one of the great sins of programming.
+In particular, large LOC of a researcher may simply mean that they write inefficient and repetitive code, one of the great sins of programming.
 A feature of good software is modularity because it enable reuse of functions both inside and outside the project.
@@
 Software peer-review is increasingly becoming recognized as a means to enable better research software.
-Peer-review can happen in the context of dedicated software journal (e.g, Journal of Statistical Software), as part of the standard review process (for guidelines, see https://www.carlboettiger.info/2013/06/13/what-I-look-for-in-software-papers.html), or as part of submissions to software archives that guarantuee certain standards (e.g., rOpenSci).
-We suggest that active participation in the review of scientific software, independent of the involvement or maintance of specific software products, should be a dimension of positive evaluation.
+Peer-review can happen in the context of dedicated software journal (e.g, Journal of Statistical Software), as part of the standard review process (for guidelines, see https://www.carlboettiger.info/2013/06/13/what-I-look-for-in-software-papers.html), or as part of submissions to software archives that guarantee certain standards (e.g., rOpenSci).
+We suggest that active participation in the review of scientific software, independent of the involvement or maintenance of specific software products, should be a dimension of positive evaluation.
 
@@
 
-In sum, we are thankful to Gaertner et al.
-for highlighting the importance of research software contributions as academically valuable products.
-Specifically, we value ...
-However, we are critical of ...
-for the reasons mentioned here.
-We propose to ...
+In sum, we are thankful to  @gartner2022responsible for highlighting the importance of research software contributions as scientifically valuable products.
+Specifically, we value the distinction of different classes of software projects (e.g., small scripts vs large-scale team efforts), ....
+However, we are critical of the proposed quantitative impact measures.
+For the reasons mentioned here, We propose to emphasize that qualitative statement on impact, in which software authors should argue in how far their project paved the way for other scientific endeavours.
 

### 2023-01-23 manuscript.Rmd
added FAIR4RS

@@
     address       : "Postal address"
-    email         : "my@email.com"
+    email         : "andreas.brandmaier@medicalschool-berlin.de"
     role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
@@
 At the same time, they are usually unrecognized because they are invisible to the end user and provide no functionality on its own.
+A specific set of guidelines implements FAIR for research software and may serve to quantify the FAIRness of research software [@barker2022introducing].
 
@@
 
-We discourage the lines of code (LOC) metric that was proposed to measure effort.
+To our knowledge, there is no consensus as how to measure productivity or quality of software contributions and no single metric seems to be a valid measure [e.g., @7332509].
+Specifically, we discourage the lines of code (LOC) metric that was proposed to measure effort.
 Lines of code is a common metric in software development useful to gauge the size of an entire code base or its change over time.

### 2023-02-01 manuscript.Rmd
every sentence on a new line
restructure

@@
 
+# Introduction
 Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a specific implementation of these principles to broaden the range of what constitutes research contributions.
-Gaertner et al.
-provide evaluation schemes for published manuscripts, published datasets, and software contributions.
+Gaertner et al. provide evaluation schemes for published manuscripts, published datasets, and software contributions.
 We specifically endorse the increased emphasis on research software as recognizable and laudable scientific contribution.
-Why are research software contributions important scientific contributions? We would like to respond with a quote from the Science Code Manifesto (http://sciencecodemanifesto.org/):
-"Software is an essential research product, and the effort to produce, maintain, adapt, and 
-curate code must be recognized."
+Why are research software contributions important scientific contributions?
+We would like to respond with a quote from the Science Code Manifesto (http://sciencecodemanifesto.org/):
+"Software is an essential research product, and the effort to produce, maintain, adapt, and curate code must be recognized."
 and
@@
 Here, we specifically comment on the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible).
-We comment on the dimensions metrics, lines of code, software engineering, impact, peer-review, usability.
+While the authors criticize the use of invalid metrics to evaluate publications (Journal Impact Factors, Citations, H-Index), those same metrics (or variants thereof) are mainly relied on in their proposal to evaluate research software.
+However, they same critisism applies to those metrics in the context of research software evaluation.
+We therefore argue to follow their evaluation scheme for publications: In a first stage, rigor should be evaluated, establishing whether the software product passes a certain treshold, and in the second stage, more complicated dimensions, such as impact, can be evaluated.
+In the following, we discuss their proposed criteria.
+We then discuss rigor in the context of research software, and propose specific criteria to measure it.
+At last, we elaborate on the assessment of impact in the context of research software.
 
-## Metrics
+# Proposed Criteria
+## ID 9: Citations
+@gartner2022responsible suggest the number of citations (their indicator #9 in Table 3) to evaluate the impact of a software package.
+However, citations suffer from serious shortcoming when used to evaluate the impact of research software.
+For example, it is much more commonly accepted to cite data analysis packages than supporting packages (such as papaja [@R-papaja], RMarkdown, RStudio, and LaTeX used to render this article).
+Therefore, the total citation count is confounded with the type of software.
+A further complication arises through the fact that functionality of successful, modular scientific software is imported, re-used, and re-exported in other software packages.
+While we highly encourage this from both a software engineering perspective and for scientific progress, it poses further difficulties to citation-based metrics.
+This especially punishes highly relied-on backends such as optimization algorithms: For example, the NLopt optimization suite is cited 1,711 times on google scholar.
+However, the algorithms from that suite are used in countless software packages, for example, as a backend of OpenMx which again has 811 citations.
+Moreover, the algorithms from that suite are all open source, and therefore serve as a reference for others.
+As an example, consider the lme4 package, that adapted one of the algortihms from NLopt.
+This package has 57,820 citations on google scholar, yet no single lme4 user probably ever cited the NLopt library.
+We therefore argue that merrit statements are much more useful to assess impact of research software.
 
-@gartner2022responsible suggest the number of citations (their indicator #9 in Table 3) to evaluate the impact of a software package.
-They acknowledge the problem that citations inconsistently point to software packages and/or accompanying software packages.
-For example, one of the authors maintains the pdc package for R.
-Google scholar counts X citations to the package and X citations to the accompanying paper, of which XX citations cite both, which permits a simple addition of the citation counts.
-
-<!-- This needs to be shortened a lot-->
-Date of first fully functional public version is a central field in the proposal as it serves as a denominator in computing citations relative to software age.
-In our own experience, research software often evolves very differently to the standards that software engineering prescribes.
-In best-practice software development, the first step is software specification or requirements engineering.
-This is the process of understanding and defining what features and services are required from the product and how they are to be offered.
-This involves research on existing solutions, discussions with users, user interface testing with mockups or user observations while testing prototypes.
-A requirement specification documents the expected functionality, such that development time and cost can be estimated and requirements can be formally checked with the final product for consistency and completeness.
-From our own experience, hardly any of these steps apply to scientific software development.
-Scientific software is often developed as a side-project fuelled by enthusiasm instead of proper funding.
-Some pieces of software start as testbed implementation of a newly proposed methodology, others come to life as wrappers or graphical interfaces for other packages, some others were never meant to be shared but are so useful that they became a package on their own.
-
+## ID 11: GitHub Stars
 We strongly object to the metric Github stars.
@@
 
-In our opinion, total citation metrics, number of users, downloads per month and similar may provide a coarse measure for the impact of a software package, even though it is important to note that these metrics are difficult to compare across different types of software packages.
-For example, it is probably more commonly accepted to cite data analysis packages (particularly if they are commercial packages) than supporting packages (such as papaja [@R-papaja] used to render this article).
-A further complication arises through the fact that functionality of successful, modular scientific software is imported, re-used, and re-exported in other software packages.
-While we highly encourage this from both a software engineering perspective and for scientific progress, it poses further difficulties to citation-based metrics.
-For example, the pdc package [@pdc-paper] offers functions to cluster time-series based on relative complexity.
-The TSclust package [@TSclust] is a wrapper package that unites several divergence measures for time series clustering conveniently in a single package, particularly, the pdc clustering approach.
-This potentially leads to researchers only citing the TSclust package but not the original algorithm (e.g., as in the work by @staroverova).
+## ID 14: Lines of Code
+We discourage the lines of code (LOC) metric that was proposed to measure effort.
+There are several problems in using this indicator.
+First of all, lines of code for code contributions highly depends on the programming language, mastery and personal programming style.
+In particular, large LOC of a researcher may simply mean that they write inefficient and repetitive code, one of the great sins of programming.
+A feature of good software is modularity because it enables reuse of functions both inside and outside the project.
+A great example is the CRAN library that enables packages to built upon other packages and reuse their functionality.
+However, if LOC were a metric that a researcher are compared with, it encourages researchers to copy&paste chunks of code from other packages instead of importing them in their project.
+Second, lines of code is an ambiguous metric.
+It can be calculated differently depending on whether comments do count or whether physical lines of code or logical lines of code are counted.
 
-Last, we would like to point out that the request for citation metrics with respect to software packages seems inconsistent with the authors condemnation of citation metrics for publications.
+## ID 5: Date of first fully functional public version
+Date of first fully functional public version is a central field in the proposal as it serves as a denominator in computing citations relative to software age.
+In our own experience, research software often evolves very differently to the standards that software engineering prescribes.
+<!-- I deleted the part about how software should be planned and written - I found it very good and interesting, but we don't have much space :/ -->
+It is often developed as a side-project fuelled by enthusiasm instead of proper funding.
+Some pieces of software start as testbed implementation of a newly proposed methodology, others come to life as wrappers or graphical interfaces for other packages, some others were never meant to be shared but are so useful that they became a package on their own.
+<!-- Missing: Whats the problem? -->
+
+## Open Source
+is not used in phase 1
+
+## ID 17: Reusability Indicator
+This is one of only two criteria used in phase one of their proposed assessment procedure, and therefore is of central importance.
+We think this item tries to assess important aspects of rigor in software development: documentation, maintenance, testing and bug chasing, for example.
+However, it conflates *usablity* with *usage*, as it also tries to assess the number of users.
+In addition, the criteria for the different categories are not sharply defined.
+For example, the current definitions read
+
+> R2 (2 points): Well-developed and tested software, fairly extensive documentation, ...
+
+> R3 (4 points): ...extensive documentation, systematic bug chasing and unit testing, ...
+
+We think that well-tested software implies systematic bug chasing and unit tests, and that the difference between "fairly extensive" and "extensive" documentation is unclear.
+As a result, this indicator is more of a "gut-feeling" indicator, roughly assessing the "size" of the software project.
+We instead propose to assess the different aspects of rigor separately, more sharply defined, and also assess "usage" indepentently from quality.
+
+# Aspects of rigor for Software Implementations
+Best practices of SE:
+## unit testing
+## CI
+## documentation
+## open source
+## bug reporting and disclosure in changelogs
+## activate maintenance
+## reproducibility
+
+# Measuring Impact
+In our opinion, total citation metrics, number of users, downloads per month and similar may provide a coarse measure for the impact of a software package, even though it is important to note the severe shortcoming we described above.
 
@@
 We believe this is a fair assessment of the real impact no matter whether someone contributed documentation/tutorial to run certain types of models, whether someone contributed a specific feature that enabled a given research project to run their analysis, or whether someone is the author of the entire package.
-
-## Software Engineering standards
-
-As mentioned before, software engineering standards are often not adhered to in the ad-hoc
-development of scientific software.
-Thus, at the minimum, we would like to encourage researchers if they provide unit tests.
-Unit tests are the backbone of software quality and testing.
-Developing unit tests may often cost more time than developing the actual code.
-At the same time, they are usually unrecognized because they are invisible to the end user and provide no functionality on its own.
-A specific set of guidelines implements FAIR for research software and may serve to quantify the FAIRness of research software [@barker2022introducing].
-
-## Lines of code
-
-To our knowledge, there is no consensus as how to measure productivity or quality of software contributions and no single metric seems to be a valid measure [e.g., @7332509].
-Specifically, we discourage the lines of code (LOC) metric that was proposed to measure effort.
-Lines of code is a common metric in software development useful to gauge the size of an entire code base or its change over time.
-There are several problems in using this indicator for effort.
-First of all, lines of code is an ambiguous metric.
-It can be calculated differently depending on whether comments do count or whether physical lines of code or logical lines of code are counted.
-Second, lines of codes for code contributions highly depends on the programming language, mastery and personal programming style.
-In particular, large LOC of a researcher may simply mean that they write inefficient and repetitive code, one of the great sins of programming.
-A feature of good software is modularity because it enable reuse of functions both inside and outside the project.
-A great example is the CRAN library that enables packages to built upon other packages and reuse their functionality.
-However, if LOC were a metric that a researcher are compared with, it encourages researchers to copy&paste chunks of code from other packages instead of importing them in their project.
-
-## Peer-Review
-
-Similar to peer-review for manuscripts, software code is amenable to peer-review. It is common
-practice in corporate software development, to have peers review code.
-For example, if a code base is changed, it is good practice to request a change with the developer team before actually changing the code base.
-The change is reviewed by experts and/or maintainers before it is integrated in the active branch of development.
-In developing these guidelines, we see the opportunity to foster such good practices and reward researchers that actively engage in reviewing computer code.
-Software peer-review is increasingly becoming recognized as a means to enable better research software.
-Peer-review can happen in the context of dedicated software journal (e.g, Journal of Statistical Software), as part of the standard review process (for guidelines, see https://www.carlboettiger.info/2013/06/13/what-I-look-for-in-software-papers.html), or as part of submissions to software archives that guarantee certain standards (e.g., rOpenSci).
-We suggest that active participation in the review of scientific software, independent of the involvement or maintenance of specific software products, should be a dimension of positive evaluation.
-
-## Reproducibility
-
-->Aaron and Max
-
-## Summary
-
-In sum, we are thankful to  @gartner2022responsible for highlighting the importance of research software contributions as scientifically valuable products.
-Specifically, we value the distinction of different classes of software projects (e.g., small scripts vs large-scale team efforts), ....
-However, we are critical of the proposed quantitative impact measures.
-For the reasons mentioned here, We propose to emphasize that qualitative statement on impact, in which software authors should argue in how far their project paved the way for other scientific endeavours.
 

### 2023-02-01 manuscript.Rmd
add a lot

@@
 # Proposed Criteria
+
 ## ID 9: Citations
@@
 
-## Open Source
-is not used in phase 1
+## ID 7: Contributer Roles and Involvement
+<!-- Larger software project, but the same effort of the individual researcher leads to lower scores? -->
+
+## ID 8: License
+Currently, whether the software is open source is not used in computing a treshold for phase one of the assessment procedure.
+However, we think the license is central for the value of  software as a research product.
+Above, we already discussed the different ways research software can have an impact - not only by direct usage, but also by reimporting and reusing software in other packages.
+This type of impact is not possible for closed-source software.
+In addition, many aspects of rigor are not possible to evaluate for closed-source software - for example, whether it is well-tested or whether bugs have been fixed.
+We therefore propose to either score the license directly, giving points to open-source licenses as defined by the Open Source Society (https://opensource.org/licenses), or to assess the indicators of rigor as discussed in the second part of this commentary, where it is only possible to gain points for certain aspects of rigor if the software is open-source.
 
@@
 As a result, this indicator is more of a "gut-feeling" indicator, roughly assessing the "size" of the software project.
-We instead propose to assess the different aspects of rigor separately, more sharply defined, and also assess "usage" indepentently from quality.
+We instead propose to assess "usage" or impact indepentently from quality/rigor.
+We also propose to assess the different aspects of rigor (testing, documentation, etc.) separately, and more sharply defined.
+We therefore dicuss in the following aspects of rigor for research software, and how they could be assessed.
 
 # Aspects of rigor for Software Implementations
-Best practices of SE:
-## unit testing
+Best-practises for software development can be used to assess rigor for research software.
+In the following, we discuss aspects that should be scored in phase one of the assessment.
+<!-- some introductory text missing -->
+<!-- maybe cite
+https://joss.readthedocs.io/en/latest/review_criteria.html
+and
+https://joss.readthedocs.io/en/latest/review_checklist.html
+here -->
+## Tests
+Tests are essential to verify the correctness of scientific software.
+They are usually separated into different categories:
+*Unit tests* evaluate only small pieces of code, for example specific R-functions.
+*Functional Tests* test whether the input of a user interacting with the software leads to the correct output.
+There are many more types of tests, and the type of tests appropriate for a specific software product depends on the software itself.
+However, we argue it should at least be tested whether the software can be installed on different platforms, and whether the core functionality of the software is executed without errors (*Smoke Test*).
+In addition, it might be possible to give higher scores if different types of tests are employed, or to define a minimum of tests types that should be present, for example *unit tests*, *integration tests* and *functional tests*.
+Another metric that could be assessed is code coverage, defined as the percentage of code lines that are executed during testing.
+However, we would like to note that this metric is to be taken with a grain of salt, as the quality and number of the tests is not taken into account.
+
 ## CI
-## documentation
+## Documentation
+
 ## open source

### 2023-02-01 manuscript.Rmd
finish restructuring

@@
 
+## ID 6: Date of most recent substantive update
+<!-- see issue #7 -->
+
 ## ID 14: Lines of Code
@@
 In the following, we discuss aspects that should be scored in phase one of the assessment.
+For the reasons outline above, we propose to score those instead of the reusability indicator (item 17 in table 3).
 <!-- some introductory text missing -->
@@
 However, we would like to note that this metric is to be taken with a grain of salt, as the quality and number of the tests is not taken into account.
+Tests should be automated, or at least reproducible.
+This rules out closed-source software tests.
 
-## CI
 ## Documentation
+Just like for tests, there are different types of documentation.
+For example, there are tutorials showcasing the usage of software with examples, and there is API documentation (e.g., docstrings) for individuall functions and classes.
+We propose to identify relevant categories of documentation for research software and score the presence of each of them separately.
+<!-- maybe propose the ones from joss: Installation Instructions, Tutorials, API, Community Guidelines -->
+Another aspect of documentation is the possibility to ask questions to the package authors (see for example the lavaan mailing list).
 
-## open source
-## bug reporting and disclosure in changelogs
-## activate maintenance
-## reproducibility
+## Maintenance
+Maintaining a software package is often more work than writing it.
+This should be reflected in the assessment procedure.
+We propose to score several aspects of maintenance seperately, such as bug fixing, changelogs, and the possibility to report bugs as a user (for example via issues).
+Note that some aspects of maintenance might not be assessed for closed-source software, for example, if changelogs do not include fixed bugs.
+
+## Peer-Review
+
+Similar to peer-review for manuscripts, software code is amenable to peer-review.
+It is common practice in software development to have peers review code.
+For example, if a code base is changed, it is good practice to request a change with the developer team before actually changing the code base.
+The change is reviewed by experts and/or maintainers before it is integrated in the active branch of development.
+In developing these guidelines, we see the opportunity to foster such good practices and reward researchers that actively engage in reviewing computer code.
+Software peer-review is increasingly becoming recognized as a means to enable better research software.
+Peer-review can happen in the context of dedicated software journal (e.g, Journal of Statistical Software), as part of the standard review process (for guidelines, see https://www.carlboettiger.info/2013/06/13/what-I-look-for-in-software-papers.html), or as part of submissions to software archives that guarantee certain standards (e.g., rOpenSci).
+We suggest that active participation in the review of scientific software, independent of the involvement or maintenance of specific software products, should be a dimension of positive evaluation.
+
+## Reproducibility
 
@@
 
+# Summary
+
+In sum, we are thankful to  @gartner2022responsible for highlighting the importance of research software contributions as scientifically valuable products.
+Specifically, we value the distinction of different classes of software projects (e.g., small scripts vs large-scale team efforts), ....
+However, we are critical of the proposed quantitative impact measures.
+For the reasons mentioned here, We propose to emphasize that qualitative statement on impact, in which software authors should argue in how far their project paved the way for other scientific endeavours.
+
 \newpage

### 2023-02-01 manuscript.Rmd
shorten
typo

@@
 here -->
+
 ## Tests
@@
 ## Peer-Review
-
 Similar to peer-review for manuscripts, software code is amenable to peer-review.

### 2023-02-14 manuscript.Rmd
Update manuscript.Rmd
Co-authored-by: Aaron Peikert <aaron.peikert@posteo.de>

@@
 # Introduction
-Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a specific implementation of these principles to broaden the range of what constitutes research contributions.
+Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a specific evaluation scheme for published manuscripts, reusable datasets, and research software.
 Gaertner et al. provide evaluation schemes for published manuscripts, published datasets, and software contributions.

### 2023-02-14 manuscript.Rmd
Update manuscript.Rmd
Co-authored-by: Aaron Peikert <aaron.peikert@posteo.de>
Update manuscript.Rmd

@@
 Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a specific evaluation scheme for published manuscripts, reusable datasets, and research software.
-Gaertner et al. provide evaluation schemes for published manuscripts, published datasets, and software contributions.
 We specifically endorse the increased emphasis on research software as recognizable and laudable scientific contribution.
@@
 We would like to respond with a quote from the Science Code Manifesto (http://sciencecodemanifesto.org/):
+"Software is a cornerstone of science. Without software, twenty-first century science would be impossible."
+and
 "Software is an essential research product, and the effort to produce, maintain, adapt, and curate code must be recognized."
-and
-"Software is a cornerstone of science. Without software, twenty-first century science would be impossible."
 Yet, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivise software development and, specifically, good software engineering practice that we would expect from a research software engineer [@baxter2012research].

### 2023-02-14 manuscript.Rmd
Apply suggestions from code review
Co-authored-by: Aaron Peikert <aaron.peikert@posteo.de>

@@
 Here, we specifically comment on the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible).
-While the authors criticize the use of invalid metrics to evaluate publications (Journal Impact Factors, Citations, H-Index), those same metrics (or variants thereof) are mainly relied on in their proposal to evaluate research software.
+While the authors criticize the use of invalid metrics to evaluate publications (Journal Impact Factors, Citations, H-Index), those same metrics (or ad hoc variants thereof) are mainly relied on in their proposal to evaluate research software.
 However, they same critisism applies to those metrics in the context of research software evaluation.
@@
 ## ID 9: Citations
+
 @gartner2022responsible suggest the number of citations (their indicator #9 in Table 3) to evaluate the impact of a software package.
@@
 ## ID 11: GitHub Stars
-We strongly object to the metric Github stars.
+
+We object to the use of Github stars as a general metric.
 First of all, it is tied only to a single, commercial platform and allows no freedom to choose free alternatives such as local Gitlab instances or other platforms.
@@
 ## ID 14: Lines of Code
+
 We discourage the lines of code (LOC) metric that was proposed to measure effort.
@@
 ## ID 5: Date of first fully functional public version
-Date of first fully functional public version is a central field in the proposal as it serves as a denominator in computing citations relative to software age.
-In our own experience, research software often evolves very differently to the standards that software engineering prescribes.
+
+The date of the first fully functional public version is a central field in the proposal as it serves as a denominator in computing citations relative to the software age.
+Besides the problem of creating an ad hoc citation metric, this date is difficult to determine and to verify.
+In our own experience, research software often evolves driven by time constraints and the need for specific functionality, not according to feature completeness.
 <!-- I deleted the part about how software should be planned and written - I found it very good and interesting, but we don't have much space :/ -->
@@
 Some pieces of software start as testbed implementation of a newly proposed methodology, others come to life as wrappers or graphical interfaces for other packages, some others were never meant to be shared but are so useful that they became a package on their own.
-<!-- Missing: Whats the problem? -->
+Feature completeness is not a valid measure in these cases.
 
@@
 However, we think the license is central for the value of  software as a research product.
-Above, we already discussed the different ways research software can have an impact - not only by direct usage, but also by reimporting and reusing software in other packages.
+Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by reimporting and reusing software in other packages.
 This type of impact is not possible for closed-source software.
-In addition, many aspects of rigor are not possible to evaluate for closed-source software - for example, whether it is well-tested or whether bugs have been fixed.
+In addition, many aspects of rigor are not possible to evaluate for closed-source software --- for example, whether it is well-tested or whether bugs have been fixed.
 We therefore propose to either score the license directly, giving points to open-source licenses as defined by the Open Source Society (https://opensource.org/licenses), or to assess the indicators of rigor as discussed in the second part of this commentary, where it is only possible to gain points for certain aspects of rigor if the software is open-source.
@@
 ## ID 17: Reusability Indicator
+
 This is one of only two criteria used in phase one of their proposed assessment procedure, and therefore is of central importance.
@@
 # Aspects of rigor for Software Implementations
+
 Best-practises for software development can be used to assess rigor for research software.
@@
 ## Tests
+
 Tests are essential to verify the correctness of scientific software.
@@
 ## Documentation
+
 Just like for tests, there are different types of documentation.
@@
 ## Maintenance
+
 Maintaining a software package is often more work than writing it.
@@
 ## Peer-Review
+
 Similar to peer-review for manuscripts, software code is amenable to peer-review.
@@
 # Measuring Impact
+
 In our opinion, total citation metrics, number of users, downloads per month and similar may provide a coarse measure for the impact of a software package, even though it is important to note the severe shortcoming we described above.
@@
 Complementing the quantitative aspect of impact, we suggest to introduce a mandatory qualitative judgement on the impact of the software, which fits with what @gartner2022responsible proposed as indicator "merit / impact statement".
-In our view, researchers should be requested to indicate at least one (and up to three) empirical research projects that directly benefited from the software contributions they made.
+In our view, researchers should be requested to indicate at least one (and up to three)  research projects that directly benefited from the software contributions they made.
 We believe this is a fair assessment of the real impact no matter whether someone contributed documentation/tutorial to run certain types of models, whether someone contributed a specific feature that enabled a given research project to run their analysis, or whether someone is the author of the entire package.

### 2023-02-17 manuscript.Rmd
updated affiliations some minor changes and shortening

@@
     corresponding : yes    # Define only one corresponding author
-    address       : "Postal address"
+    address       : "Rüdesheimer Str. 50, 14197 Berlin"
     email         : "andreas.brandmaier@medicalschool-berlin.de"
@@
   - name          : "Maximilian Ernst"
-    affiliation   : "2"
+    affiliation   : "2,4"
     role:
@@
   - name          : "Aaron Peikert"
-    affiliation   : "2"
+    affiliation   : "2,3,4"
     role:
@@
   - id            : "3"
+    institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research"
+  - id            : "4"
     institution   : "Humboldt Universität zu Berlin"
@@
 Yet, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivise software development and, specifically, good software engineering practice that we would expect from a research software engineer [@baxter2012research].
-Here, we specifically comment on the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible).
-While the authors criticize the use of invalid metrics to evaluate publications (Journal Impact Factors, Citations, H-Index), those same metrics (or ad hoc variants thereof) are mainly relied on in their proposal to evaluate research software.
+We suggest that criteria for research software contributions must capture two major dimensions: rigor and impact.
+Rigor means implementing highest standards and best practices for ensuring transparency, correctness and reusability of a software.
+From this perspective, we comment on the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible) and propose alternatives to better capture rigor and impact.
+
+<!--While the authors criticize the use of invalid metrics to evaluate publications (Journal Impact Factors, Citations, H-Index), those same metrics (or ad hoc variants thereof) are mainly relied on in their proposal to evaluate research software.
 However, they same critisism applies to those metrics in the context of research software evaluation.
@@
 At last, we elaborate on the assessment of impact in the context of research software.
+
+-->
 
@@
 This package has 57,820 citations on google scholar, yet no single lme4 user probably ever cited the NLopt library.
-We therefore argue that merrit statements are much more useful to assess impact of research software.
+We therefore argue that qualitative merrit statements are much more useful to assess impact of research software.
 
@@
 We object to the use of Github stars as a general metric.
-First of all, it is tied only to a single, commercial platform and allows no freedom to choose free alternatives such as local Gitlab instances or other platforms.
-More importantly, we question the validity of such metrics that like rating systems on commercial platforms are too easy to be gamed by creating fake accounts that provide high ratings.
+First of all, it is tied only to a single, commercial platform.
+More importantly, we question the validity of such metrics that, like rating systems on commercial platforms, are too easy to be gamed.
 
@@
 However, if LOC were a metric that a researcher are compared with, it encourages researchers to copy&paste chunks of code from other packages instead of importing them in their project.
-Second, lines of code is an ambiguous metric.
-It can be calculated differently depending on whether comments do count or whether physical lines of code or logical lines of code are counted.
+<!--Second, lines of code is an ambiguous metric.
+It can be calculated differently depending on whether comments do count or whether physical lines of code or logical lines of code are counted.-->
 
@@
 ## ID 8: License
-Currently, whether the software is open source is not used in computing a treshold for phase one of the assessment procedure.
-However, we think the license is central for the value of  software as a research product.
-Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by reimporting and reusing software in other packages.
-This type of impact is not possible for closed-source software.
+At present, the question of whether the software is open source is not used in the computation of a treshold for phase one of the assessment procedure.
+However, we think an open license is central for assessing both rigor and (potential) impact.
+Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by re-using software in other packages.
+Broader impact in science is more likely with open source software.
 In addition, many aspects of rigor are not possible to evaluate for closed-source software --- for example, whether it is well-tested or whether bugs have been fixed.
-We therefore propose to either score the license directly, giving points to open-source licenses as defined by the Open Source Society (https://opensource.org/licenses), or to assess the indicators of rigor as discussed in the second part of this commentary, where it is only possible to gain points for certain aspects of rigor if the software is open-source.
+Therefore, we propose to either score a package's license, giving points to open-source licenses (those approved in a review process by the Open Source Society (https://opensource.org/licenses).
+<!--Alternatively, one may indirectly score openness by assessing the indicators of rigor as discussed in the second part of this commentary, where it is only possible to gain points for certain aspects of rigor if the software is open-source.-->
 
@@
 As a result, this indicator is more of a "gut-feeling" indicator, roughly assessing the "size" of the software project.
-We instead propose to assess "usage" or impact indepentently from quality/rigor.
+We instead propose to assess "usage" or impact independently of rigor.
 We also propose to assess the different aspects of rigor (testing, documentation, etc.) separately, and more sharply defined.
@@
 
-## Reproducibility
-
 # Measuring Impact

### 2023-02-20 manuscript.Rmd
minor restructuring some rewriting and shortening removed peer review part added front note

@@
 ---
-title             : "A commentary to 'Responsible Research Assessment II: A specific proposal for hiring and promotion in psychology'"
+title             : "Commentary: 'Responsible Research Assessment II: A specific proposal for hiring and promotion in psychology'"
 shorttitle        : "Responsible research assessment"
@@
 
+note: |
+  A commentary on:
+  Gärtner, A., Leising, D., & Schönbrodt, F.
+D.
+(2022, November 25).
+Responsible Research Assessment II: A specific proposal for hiring and promotion in psychology.
+https://doi.org/10.31234/osf.io/5yexm
+
+  To be submitted to: Meta-Psychology
+
 affiliation:
@@
 
-abstract: |
-  TBD
+correspondence: |
+  brandmaier@mpib-berlin.mpg.de
 
-keywords          : "keywords"
-wordcount         : "X"
+#abstract: |
+#  TBD
+
+keywords          : "research software, open science, metrics, rigor, impact"
+wordcount         : "`r tryCatch(wordcountaddin::word_count(here::here('manuscript.Rmd')))`"
 
@@
 floatsintext      : no
-linenumbers       : yes
+linenumbers       : no
 draft             : no
@@
 output            : papaja::apa6_pdf
+
+equal_contrib: yes
+
+header-includes: |
+  \usepackage{hyperref}
 ---
@@
 We suggest that criteria for research software contributions must capture two major dimensions: rigor and impact.
-Rigor means implementing highest standards and best practices for ensuring transparency, correctness and reusability of a software.
+Impact measures whether the scholarly effort and the implementation and dissemination actually had an effect on the field.  
+Rigor means implementing highest standards and best practices for ensuring transparency, correctness and reusability of a piece of software. 
+By setting a high bar of rigor in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and thus better science.
 From this perspective, we comment on the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible) and propose alternatives to better capture rigor and impact.
@@
 
+In the following, we comment on individual indicators proposed in Table 3 [@gartner2022responsible].
+
+## ID 5: Date of first fully functional public version
+
+The date of the first fully functional public version serves as a denominator in computing the number of citations (see ID 9) relative to software age.
+In our experience, research software often evolves driven by time constraints and the need for specific functionality, not according to feature completeness. 
+Thus, it may generally suffice to ask for the age of the software relative to its first public release without any constraints on functionality.
+
+## ID 6: Date of most recent substantive update
+
+We believe that it's not always clear when updates are considered 'substantive'. 
+As an important aspect of rigor, we believe that software needs to be actively maintained.
+To assess this, we propose to simply provide a checkbox, in which the author indicates whether a scientific software package is actively maintained (e.g., software has a regular release cycle or an update in the last six months).
+In addition, authors have a chance to explain why their active maintenance may be different from these guidelines.
+
 ## ID 9: Citations
 
-@gartner2022responsible suggest the number of citations (their indicator #9 in Table 3) to evaluate the impact of a software package.
-However, citations suffer from serious shortcoming when used to evaluate the impact of research software.
-For example, it is much more commonly accepted to cite data analysis packages than supporting packages (such as papaja [@R-papaja], RMarkdown, RStudio, and LaTeX used to render this article).
-Therefore, the total citation count is confounded with the type of software.
-A further complication arises through the fact that functionality of successful, modular scientific software is imported, re-used, and re-exported in other software packages.
-While we highly encourage this from both a software engineering perspective and for scientific progress, it poses further difficulties to citation-based metrics.
-This especially punishes highly relied-on backends such as optimization algorithms: For example, the NLopt optimization suite is cited 1,711 times on google scholar.
-However, the algorithms from that suite are used in countless software packages, for example, as a backend of OpenMx which again has 811 citations.
-Moreover, the algorithms from that suite are all open source, and therefore serve as a reference for others.
-As an example, consider the lme4 package, that adapted one of the algortihms from NLopt.
-This package has 57,820 citations on google scholar, yet no single lme4 user probably ever cited the NLopt library.
-We therefore argue that qualitative merrit statements are much more useful to assess impact of research software.
+@gartner2022responsible suggest the number of citations to evaluate the impact of a software package.
+However, citations suffer from serious shortcomings when used to evaluate the impact of research software, particularly, when used to compare candidates in the context of hire and promotion.
+For example, it is much more commonly accepted to cite data analysis packages than supporting packages (such as papaja [@R-papaja] used to render this article). 
+Further, functionality of successful, modular scientific software is ideally imported from other software packages to avoid duplication of code and enable faster development of new software.
+While we highly encourage this from both a software engineering perspective and for scientific progress, it challenges the validity of citation-based metrics for impact.
+For example, consider the NLopt optimization suite, which counts 1,711 citations on Google scholar at the time of writing.
+NLopt is a backbone for many scientific model fitting packages both because it implements various optimization algorithm but also because it is open-source and can inspire re-implementations of these algorithms.
+For example, consider the famous lme4 package [@bates2014fitting] for generalized linear mixed-effects models, that is partly based on NLopt.
+This package has more than 58,000 citations on Google scholar, yet it is unlikely that researchers will cite the underlying optimization algorithm.
 
@@
 
-## ID 6: Date of most recent substantive update
-<!-- see issue #7 -->
-
 ## ID 14: Lines of Code
 
-We discourage the lines of code (LOC) metric that was proposed to measure effort.
+To measure effort, we discourage the lines of code (LOC) metric.
 There are several problems in using this indicator.
@@
 A feature of good software is modularity because it enables reuse of functions both inside and outside the project.
-A great example is the CRAN library that enables packages to built upon other packages and reuse their functionality.
+<!--A great example is the CRAN library that enables packages to built upon other packages and reuse their functionality.
 However, if LOC were a metric that a researcher are compared with, it encourages researchers to copy&paste chunks of code from other packages instead of importing them in their project.
@@
 It can be calculated differently depending on whether comments do count or whether physical lines of code or logical lines of code are counted.-->
-
-## ID 5: Date of first fully functional public version
-
-The date of the first fully functional public version is a central field in the proposal as it serves as a denominator in computing citations relative to the software age.
-Besides the problem of creating an ad hoc citation metric, this date is difficult to determine and to verify.
-In our own experience, research software often evolves driven by time constraints and the need for specific functionality, not according to feature completeness.
-<!-- I deleted the part about how software should be planned and written - I found it very good and interesting, but we don't have much space :/ -->
-It is often developed as a side-project fuelled by enthusiasm instead of proper funding.
-Some pieces of software start as testbed implementation of a newly proposed methodology, others come to life as wrappers or graphical interfaces for other packages, some others were never meant to be shared but are so useful that they became a package on their own.
-Feature completeness is not a valid measure in these cases.
 
@@
 ## ID 8: License
-At present, the question of whether the software is open source is not used in the computation of a treshold for phase one of the assessment procedure.
-However, we think an open license is central for assessing both rigor and (potential) impact.
+At present, the question of whether the software is open source is not used in the computation of a threshold in the prescreening phase.
+However, we think an open license is central for assessing both rigor and (potentially) increasing impact.
 Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by re-using software in other packages.
@@
 In addition, many aspects of rigor are not possible to evaluate for closed-source software --- for example, whether it is well-tested or whether bugs have been fixed.
-Therefore, we propose to either score a package's license, giving points to open-source licenses (those approved in a review process by the Open Source Society (https://opensource.org/licenses).
+Therefore, we propose to penalize software if it does not adhere to an open-source licenses (those approved in a review process by the Open Source Society (https://opensource.org/licenses) by allowing them only half of the maximum achievable total points.
 <!--Alternatively, one may indirectly score openness by assessing the indicators of rigor as discussed in the second part of this commentary, where it is only possible to gain points for certain aspects of rigor if the software is open-source.-->
@@
 
-This is one of only two criteria used in phase one of their proposed assessment procedure, and therefore is of central importance.
+This is one of only two criteria used in the prescreening phase of the proposal, and therefore is of central importance.
 We think this item tries to assess important aspects of rigor in software development: documentation, maintenance, testing and bug chasing, for example.
-However, it conflates *usablity* with *usage*, as it also tries to assess the number of users.
-In addition, the criteria for the different categories are not sharply defined.
-For example, the current definitions read
+By also incorporating the size of the user base, it confounds *usablity* (as an aspect of rigor) with *usage* (as an aspect of impact).
+Importantly, the criteria for the different proposed categories are not clearly defined.
+For example, the difference between "fairly extensive" and "extensive" documentation is unclear.
+As a result, this indicator is more of a "gut-feeling" indicator, roughly assessing the "size" of the software project.
+We instead propose to assess impact independently of rigor in phase I.
+In the following, we propose ways to assess the different aspects of rigor separately, and more clearly.
 
-> R2 (2 points): Well-developed and tested software, fairly extensive documentation, ...
+# Rigor for software implementations
 
-> R3 (4 points): ...extensive documentation, systematic bug chasing and unit testing, ...
-
-We think that well-tested software implies systematic bug chasing and unit tests, and that the difference between "fairly extensive" and "extensive" documentation is unclear.
-As a result, this indicator is more of a "gut-feeling" indicator, roughly assessing the "size" of the software project.
-We instead propose to assess "usage" or impact independently of rigor.
-We also propose to assess the different aspects of rigor (testing, documentation, etc.) separately, and more sharply defined.
-We therefore dicuss in the following aspects of rigor for research software, and how they could be assessed.
-
-# Aspects of rigor for Software Implementations
-
-Best-practises for software development can be used to assess rigor for research software.
-In the following, we discuss aspects that should be scored in phase one of the assessment.
-For the reasons outline above, we propose to score those instead of the reusability indicator (item 17 in table 3).
+In the following, we discuss aspects that we believe should be scored in phase one of the assessment.
+We propose to use these as equally weighted indicators in lieu of the proposed coarse reusability indicator (item 17 in Table 3 of the proposal).
 <!-- some introductory text missing -->
@@
 
-Tests are essential to verify the correctness of scientific software.
-They are usually separated into different categories:
+Tests are essential to discover invalid or incorrect functionality, investigate scalability of code and reveal poor design choices.
+There are a variety of useful tests, such as unit tests of software subcomponents or tests of software functionality at larger scale (e.g., see the testthat package in R).
+It would be possible to quantify aspects of software testing, for example, by assessing code coverage defined as the percentage of code lines that are executed during testing.
+However, we believe that we should give a point for software that promises that major functionality is covered by tests.
+<!--They are usually separated into different categories:
 *Unit tests* evaluate only small pieces of code, for example specific R-functions.
@@
 This rules out closed-source software tests.
-
+-->
 ## Documentation
@@
 Just like for tests, there are different types of documentation.
-For example, there are tutorials showcasing the usage of software with examples, and there is API documentation (e.g., docstrings) for individuall functions and classes.
+For example, there are tutorials showcasing the usage of software with examples, and there is application programming interface (API) documentation for individual functions and classes.
 We propose to identify relevant categories of documentation for research software and score the presence of each of them separately.
 <!-- maybe propose the ones from joss: Installation Instructions, Tutorials, API, Community Guidelines -->
-Another aspect of documentation is the possibility to ask questions to the package authors (see for example the lavaan mailing list).
+<!--AB: moved this to maintenance; Another aspect of documentation is the possibility to ask questions to the package authors (see for example the lavaan mailing list).
+-->
 
@@
 This should be reflected in the assessment procedure.
-We propose to score several aspects of maintenance seperately, such as bug fixing, changelogs, and the possibility to report bugs as a user (for example via issues).
-Note that some aspects of maintenance might not be assessed for closed-source software, for example, if changelogs do not include fixed bugs.
+We propose to score two aspects of maintenance separately, such as active bug fixing, documenting changes in logs, and the possibility to report bugs as a user (for example via tickets or mailing lists).
 
-## Peer-Review
+## Measuring Impact
 
-Similar to peer-review for manuscripts, software code is amenable to peer-review.
-It is common practice in software development to have peers review code.
-For example, if a code base is changed, it is good practice to request a change with the developer team before actually changing the code base.
-The change is reviewed by experts and/or maintainers before it is integrated in the active branch of development.
-In developing these guidelines, we see the opportunity to foster such good practices and reward researchers that actively engage in reviewing computer code.
-Software peer-review is increasingly becoming recognized as a means to enable better research software.
-Peer-review can happen in the context of dedicated software journal (e.g, Journal of Statistical Software), as part of the standard review process (for guidelines, see https://www.carlboettiger.info/2013/06/13/what-I-look-for-in-software-papers.html), or as part of submissions to software archives that guarantee certain standards (e.g., rOpenSci).
-We suggest that active participation in the review of scientific software, independent of the involvement or maintenance of specific software products, should be a dimension of positive evaluation.
-
-# Measuring Impact
-
-In our opinion, total citation metrics, number of users, downloads per month and similar may provide a coarse measure for the impact of a software package, even though it is important to note the severe shortcoming we described above.
+In our opinion, total citation metrics, number of users, downloads per month and similar may provide a coarse measure for the impact of a software package, even though it is important to note the shortcomings we described above.
 
 Complementing the quantitative aspect of impact, we suggest to introduce a mandatory qualitative judgement on the impact of the software, which fits with what @gartner2022responsible proposed as indicator "merit / impact statement".
+
 In our view, researchers should be requested to indicate at least one (and up to three)  research projects that directly benefited from the software contributions they made.
 We believe this is a fair assessment of the real impact no matter whether someone contributed documentation/tutorial to run certain types of models, whether someone contributed a specific feature that enabled a given research project to run their analysis, or whether someone is the author of the entire package.
+
+We therefore argue that qualitative merit statements are much more useful to assess impact of research software and that this should be the primary statement for committee members to evaluate if they are less concerned with the technical aspects of research software development.
 
@@
 In sum, we are thankful to  @gartner2022responsible for highlighting the importance of research software contributions as scientifically valuable products.
-Specifically, we value the distinction of different classes of software projects (e.g., small scripts vs large-scale team efforts), ....
-However, we are critical of the proposed quantitative impact measures.
-For the reasons mentioned here, We propose to emphasize that qualitative statement on impact, in which software authors should argue in how far their project paved the way for other scientific endeavours.
+We believe that the current proposal should aim to better reflect the distinction between rigor and impact.
+To this end, we emphasize a more fine-grained assessment of rigor and a qualitative statement on impact, in which software authors should argue in how far their project paved the way for other scientific endeavours.
 

### 2023-02-21 manuscript.Rmd
some more changes

@@
 We suggest that criteria for research software contributions must capture two major dimensions: rigor and impact.
-Impact measures whether the scholarly effort and the implementation and dissemination actually had an effect on the field.  
+Impact measures whether the scholarly effort and the implementation and dissemination actually had a visible effect on the field.  
 Rigor means implementing highest standards and best practices for ensuring transparency, correctness and reusability of a piece of software. 
@@
 As an important aspect of rigor, we believe that software needs to be actively maintained.
-To assess this, we propose to simply provide a checkbox, in which the author indicates whether a scientific software package is actively maintained (e.g., software has a regular release cycle or an update in the last six months).
+To assess this, we propose to simply provide a checkbox, in which the author indicates whether a scientific software package is actively maintained (e.g., software has a regular release cycle or an update within the last six months).
 In addition, authors have a chance to explain why their active maintenance may be different from these guidelines.
@@
 @gartner2022responsible suggest the number of citations to evaluate the impact of a software package.
-However, citations suffer from serious shortcomings when used to evaluate the impact of research software, particularly, when used to compare candidates in the context of hire and promotion.
+However, citations suffer from serious shortcomings when used to evaluate the impact of research software, particularly, when used comparatively in the context of hire and promotion.
 For example, it is much more commonly accepted to cite data analysis packages than supporting packages (such as papaja [@R-papaja] used to render this article). 
-Further, functionality of successful, modular scientific software is ideally imported from other software packages to avoid duplication of code and enable faster development of new software.
+Further, functionality of successful, modular scientific software is ideally reused in other software packages to avoid duplication of code and enable faster development of new software.
 While we highly encourage this from both a software engineering perspective and for scientific progress, it challenges the validity of citation-based metrics for impact.
@@
 This package has more than 58,000 citations on Google scholar, yet it is unlikely that researchers will cite the underlying optimization algorithm.
-
+<!--For example, the pdc package [@pdc-paper] offers functions to cluster time-series based on relative complexity.
+As another example, the TSclust package [@TSclust] is a wrapper to several dissimilarity measures for clustering time series. 
+Some of us maintain the pdc clustering package [@pdc-paper], which is contained in TSclust and we saw that researchers use the pdc approach but only cite the TSclust package.
+-->
 ## ID 11: GitHub Stars
@@
 To measure effort, we discourage the lines of code (LOC) metric.
-There are several problems in using this indicator.
-First of all, lines of code for code contributions highly depends on the programming language, mastery and personal programming style.
+LOC highly depends on programming language, mastery and personal programming style.
 In particular, large LOC of a researcher may simply mean that they write inefficient and repetitive code, one of the great sins of programming.
 A feature of good software is modularity because it enables reuse of functions both inside and outside the project.
+
 <!--A great example is the CRAN library that enables packages to built upon other packages and reuse their functionality.
@@
 At present, the question of whether the software is open source is not used in the computation of a threshold in the prescreening phase.
-However, we think an open license is central for assessing both rigor and (potentially) increasing impact.
+However, we think an open license is central for assessing both rigor and (potentially) increasing impact and should be part of the phase I assessment.
 Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by re-using software in other packages.
@@
 In addition, many aspects of rigor are not possible to evaluate for closed-source software --- for example, whether it is well-tested or whether bugs have been fixed.
-Therefore, we propose to penalize software if it does not adhere to an open-source licenses (those approved in a review process by the Open Source Society (https://opensource.org/licenses) by allowing them only half of the maximum achievable total points.
+Therefore, we propose to penalize software if it does not adhere to an open-source licenses (those approved in a review process by the Open Source Society (https://opensource.org/licenses) by allowing them only half of the achievable total points.
 <!--Alternatively, one may indirectly score openness by assessing the indicators of rigor as discussed in the second part of this commentary, where it is only possible to gain points for certain aspects of rigor if the software is open-source.-->
@@
 This is one of only two criteria used in the prescreening phase of the proposal, and therefore is of central importance.
-We think this item tries to assess important aspects of rigor in software development: documentation, maintenance, testing and bug chasing, for example.
-By also incorporating the size of the user base, it confounds *usablity* (as an aspect of rigor) with *usage* (as an aspect of impact).
+This item is important because it assesses aspects of rigor in software development: documentation, active maintenance, and testing.
+However, by incorporating the size of the user base, it confounds *usablity* (as an aspect of rigor) with *usage* (as an aspect of impact).
 Importantly, the criteria for the different proposed categories are not clearly defined.
@@
 As a result, this indicator is more of a "gut-feeling" indicator, roughly assessing the "size" of the software project.
-We instead propose to assess impact independently of rigor in phase I.
-In the following, we propose ways to assess the different aspects of rigor separately, and more clearly.
+Instead, in the following we propose to assess rigor and impact independently as primary aspects of a software contribution.
 
@@
 
-In the following, we discuss aspects that we believe should be scored in phase one of the assessment.
-We propose to use these as equally weighted indicators in lieu of the proposed coarse reusability indicator (item 17 in Table 3 of the proposal).
+<!--In the following, we discuss aspects that we believe should be scored in phase one of the assessment.-->
+We propose to use the following aspects as equally weighted indicators of rigor in lieu of the proposed coarse reusability indicator (item 17 in Table 3 of the proposal).
 <!-- some introductory text missing -->
@@
 
-Tests are essential to discover invalid or incorrect functionality, investigate scalability of code and reveal poor design choices.
-There are a variety of useful tests, such as unit tests of software subcomponents or tests of software functionality at larger scale (e.g., see the testthat package in R).
+Tests are essential to discover incorrect functionality, investigate scalability of code and reveal poor design choices.
+There are a variety of useful tests, such as unit tests of subcomponents or tests of software functionality at larger scale (e.g., see the testthat package in R, @wickham2011).
 It would be possible to quantify aspects of software testing, for example, by assessing code coverage defined as the percentage of code lines that are executed during testing.
-However, we believe that we should give a point for software that promises that major functionality is covered by tests.
+However, we believe that we should give one point for software that promises that major functionality is covered by tests.
 <!--They are usually separated into different categories:
@@
 Just like for tests, there are different types of documentation.
-For example, there are tutorials showcasing the usage of software with examples, and there is application programming interface (API) documentation for individual functions and classes.
+For example, there are tutorials showcasing the usage of software with examples, and there is application programming interface (API) documentation for individual functions and classes to enable reuse of functions in other software packages.
 We propose to identify relevant categories of documentation for research software and score the presence of each of them separately.
@@
 This should be reflected in the assessment procedure.
-We propose to score two aspects of maintenance separately, such as active bug fixing, documenting changes in logs, and the possibility to report bugs as a user (for example via tickets or mailing lists).
+We propose to score two aspects of maintenance separately, maintaining the actual code base (such as active bug fixing, documenting changes in logs), and maintaining the community (such as providing the possibility to report bugs, feature requests, or support requests via tickets or mailing lists).
 
@@
 
-Complementing the quantitative aspect of impact, we suggest to introduce a mandatory qualitative judgement on the impact of the software, which fits with what @gartner2022responsible proposed as indicator "merit / impact statement".
+We believe that the suggested merit statement is most useful to assess impact of research software and that this should be the primary statement for committee members to evaluate if they are less concerned with the technical aspects of research software development.
+
+<!--Complementing the quantitative aspect of impact, we suggest to introduce a mandatory qualitative judgement on the impact of the software, which fits with what @gartner2022responsible proposed as indicator "merit / impact statement".-->
 
 In our view, researchers should be requested to indicate at least one (and up to three)  research projects that directly benefited from the software contributions they made.
-We believe this is a fair assessment of the real impact no matter whether someone contributed documentation/tutorial to run certain types of models, whether someone contributed a specific feature that enabled a given research project to run their analysis, or whether someone is the author of the entire package.
-
-We therefore argue that qualitative merit statements are much more useful to assess impact of research software and that this should be the primary statement for committee members to evaluate if they are less concerned with the technical aspects of research software development.
+We believe this is a fair assessment of the real impact no matter whether someone contributed documentation to run certain types of models, whether someone contributed a specific feature that enabled a given research project to use the software, or whether someone is the primary author of the entire package.
 
@@
 In sum, we are thankful to  @gartner2022responsible for highlighting the importance of research software contributions as scientifically valuable products.
-We believe that the current proposal should aim to better reflect the distinction between rigor and impact.
+We believe that the current proposal must aim to better reflect the distinction between rigor and impact.
 To this end, we emphasize a more fine-grained assessment of rigor and a qualitative statement on impact, in which software authors should argue in how far their project paved the way for other scientific endeavours.
+Last, we hope to arrive at an evaluation scheme that incentivises the development of scientific open source software.
 

### 2023-02-21 manuscript.Rmd
Update manuscript.Rmd
merged github stars into later test as suggested by @aaronpeikert

@@
 -->
-## ID 11: GitHub Stars
-
-We object to the use of Github stars as a general metric.
-First of all, it is tied only to a single, commercial platform.
-More importantly, we question the validity of such metrics that, like rating systems on commercial platforms, are too easy to be gamed.
 
@@
 
-In our opinion, total citation metrics, number of users, downloads per month and similar may provide a coarse measure for the impact of a software package, even though it is important to note the shortcomings we described above.
+In our opinion, total citation metrics, number of users, downloads per month, github stars and similar may provide a coarse measure for the impact of a software package, even though it is important to note the shortcomings we described above.
 

### 2023-02-21 manuscript.Rmd
Update manuscript.Rmd
should be OSI!

@@
 In addition, many aspects of rigor are not possible to evaluate for closed-source software --- for example, whether it is well-tested or whether bugs have been fixed.
-Therefore, we propose to penalize software if it does not adhere to an open-source licenses (those approved in a review process by the Open Source Society (https://opensource.org/licenses) by allowing them only half of the achievable total points.
+Therefore, we propose to penalize software if it does not adhere to an open-source licenses (those approved in a review process by the Open Source Initiative (https://opensource.org/licenses) by allowing them only half of the achievable total points.
 <!--Alternatively, one may indirectly score openness by assessing the indicators of rigor as discussed in the second part of this commentary, where it is only possible to gain points for certain aspects of rigor if the software is open-source.-->

### 2023-02-21 manuscript.Rmd
Update manuscript.Rmd
Co-authored-by: Aaron Peikert <aaron.peikert@posteo.de>
Update manuscript.Rmd
pointing out inconsitency with DORA now.

@@
 @gartner2022responsible suggest the number of citations to evaluate the impact of a software package.
-However, citations suffer from serious shortcomings when used to evaluate the impact of research software, particularly, when used comparatively in the context of hire and promotion.
+First, we like to point out an inconsistency of proposing citation-based metrics and the propositions of @schonbrodt2022responsible.
+who reminded us that a core principle of the implementation of DORA is "abandon the use of invalid quantitative metrics of research quality and productivity in hiring and promotion" (p.2) and instead to focus on individual achievements.
+Further, citations suffer from serious shortcomings when used to evaluate the impact of research software, particularly, when used comparatively in the context of hire and promotion.
 For example, it is much more commonly accepted to cite data analysis packages than supporting packages (such as papaja [@R-papaja] used to render this article). 
@@
 This package has more than 58,000 citations on Google scholar, yet it is unlikely that researchers will cite the underlying optimization algorithm.
-<!--For example, the pdc package [@pdc-paper] offers functions to cluster time-series based on relative complexity.
-As another example, the TSclust package [@TSclust] is a wrapper to several dissimilarity measures for clustering time series. 
+For another example, the pdc package [@pdc-paper] offers functions to cluster time-series based on relative complexity.
+<!--As another example, the TSclust package [@TSclust] is a wrapper to several dissimilarity measures for clustering time series. 
 Some of us maintain the pdc clustering package [@pdc-paper], which is contained in TSclust and we saw that researchers use the pdc approach but only cite the TSclust package.
@@
 In sum, we are thankful to  @gartner2022responsible for highlighting the importance of research software contributions as scientifically valuable products.
-We believe that the current proposal must aim to better reflect the distinction between rigor and impact.
+We believe that the current proposal must aim to better reflect the distinction between rigor and impact for software like similar to what was proposed to evaluate journal articles.
 To this end, we emphasize a more fine-grained assessment of rigor and a qualitative statement on impact, in which software authors should argue in how far their project paved the way for other scientific endeavours.

### 2023-02-22 manuscript.Rmd
improve clarity, correct grammar

--- 2023-02-21 manuscript.Rmd
@@
   - id            : "4"
-    institution   : "Humboldt Universität zu Berlin"
+    institution   : "Humboldt-Universität zu Berlin"
 
@@
 # Introduction
-Based on four basic principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a specific evaluation scheme for published manuscripts, reusable datasets, and research software.
-We specifically endorse the increased emphasis on research software as recognizable and laudable scientific contribution.
+Based on four principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a specific evaluation scheme for published manuscripts, reusable datasets, and research software.
+We specifically endorse the increased emphasis on research software as a recognizable and laudable scientific contribution.
 Why are research software contributions important scientific contributions?
@@
 "Software is an essential research product, and the effort to produce, maintain, adapt, and curate code must be recognized."
-Yet, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivise software development  and, specifically, good software engineering practice that we would expect from a research software engineer [@baxter2012research].
-We suggest that criteria for research software contributions must capture two major dimensions: rigor and impact.
-Impact measures whether the scholarly effort and the implementation and dissemination actually had a visible effect on the field.  
-Rigor means implementing highest standards and best practices for ensuring transparency, correctness and reusability of a piece of software. 
-By setting a high bar of rigor in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and thus better science.
+However, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivize software development and, specifically, good software engineering practice [@baxter2012research].
+In line with principle 3 of @schonbrodt2022responsible, we suggest that criteria for research software contributions must capture two major dimensions: rigor and impact.
+Impact measures whether the scholarly effort, implementation, and dissemination actually had a visible effect on the field.
+Rigor means implementing high standards and best practices for ensuring transparency, correctness, and reusability of a piece of software.
+By setting a high bar of rigor in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and, thus, better science.
 From this perspective, we comment on the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible) and propose alternatives to better capture rigor and impact.
 
-<!--While the authors criticize the use of invalid metrics to evaluate publications (Journal Impact Factors, Citations, H-Index), those same metrics (or ad hoc variants thereof) are mainly relied on in their proposal to evaluate research software.
-However, they same critisism applies to those metrics in the context of research software evaluation.
-We therefore argue to follow their evaluation scheme for publications: In a first stage, rigor should be evaluated, establishing whether the software product passes a certain treshold, and in the second stage, more complicated dimensions, such as impact, can be evaluated.
-In the following, we discuss their proposed criteria.
-We then discuss rigor in the context of research software, and propose specific criteria to measure it.
-At last, we elaborate on the assessment of impact in the context of research software.
-
--->
-
 # Proposed Criteria
-
-In the following, we comment on individual indicators proposed in Table 3 [@gartner2022responsible].
 
@@
 
-We believe that it's not always clear when updates are considered 'substantive'. 
-As an important aspect of rigor, we believe that software needs to be actively maintained.
-To assess this, we propose to simply provide a checkbox, in which the author indicates whether a scientific software package is actively maintained (e.g., software has a regular release cycle or an update within the last six months).
+We believe it is not always clear when updates are considered 'substantive.'
+As an important aspect of rigor, software needs to be actively maintained.
+To assess this, we propose to provide a checkbox in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
 In addition, authors have a chance to explain why their active maintenance may be different from these guidelines.
@@
 @gartner2022responsible suggest the number of citations to evaluate the impact of a software package.
-However, citations suffer from serious shortcomings when used to evaluate the impact of research software, particularly, when used comparatively in the context of hire and promotion.
-For example, it is much more commonly accepted to cite data analysis packages than supporting packages (such as papaja [@R-papaja] used to render this article). 
-Further, functionality of successful, modular scientific software is ideally reused in other software packages to avoid duplication of code and enable faster development of new software.
+First, we like to point out an inconsistency in proposing citation-based metrics and the propositions of @schonbrodt2022responsible, who reminded us that a core principle of the implementation of DORA is "abandon the use of invalid quantitative metrics of research quality and productivity in hiring and promotion" (p.2) and instead to focus on individual achievements.
+Further, citations suffer from serious shortcomings when used to evaluate the impact of research software, particularly when used comparatively in the context of hire and promotion.
+For example, citing data analysis packages is much more commonly accepted than citing supporting packages (such as papaja [@R-papaja] used to render this article).
+Further, functionality of successful, modular scientific software is ideally reused in other software packages to avoid code dublication and enable faster development of new software.
 While we highly encourage this from both a software engineering perspective and for scientific progress, it challenges the validity of citation-based metrics for impact.
 For example, consider the NLopt optimization suite, which counts 1,711 citations on Google scholar at the time of writing.
-NLopt is a backbone for many scientific model fitting packages both because it implements various optimization algorithm but also because it is open-source and can inspire re-implementations of these algorithms.
+NLopt is a backbone for many scientific packages both because it implements various optimization algorithms but also because it is open-source and can inspire re-implementations of these algorithms.
 For example, consider the famous lme4 package [@bates2014fitting] for generalized linear mixed-effects models, that is partly based on NLopt.
 This package has more than 58,000 citations on Google scholar, yet it is unlikely that researchers will cite the underlying optimization algorithm.
-<!--For example, the pdc package [@pdc-paper] offers functions to cluster time-series based on relative complexity.
-As another example, the TSclust package [@TSclust] is a wrapper to several dissimilarity measures for clustering time series. 
-Some of us maintain the pdc clustering package [@pdc-paper], which is contained in TSclust and we saw that researchers use the pdc approach but only cite the TSclust package.
--->
+For another example, the pdc package [@pdc-paper] offers functions to cluster time series based on relative complexity.
+<!-- Here is a sentence missing about why this is an example that shows the invalidity of citation based metrics -->
 
@@
 
-To measure effort, we discourage the lines of code (LOC) metric.
-LOC highly depends on programming language, mastery and personal programming style.
-In particular, large LOC of a researcher may simply mean that they write inefficient and repetitive code, one of the great sins of programming.
-A feature of good software is modularity because it enables reuse of functions both inside and outside the project.
+We discourage the lines of code (LOC) metric to measure effort.
+LOC highly depends on programming language, mastery, and personal programming style.
+In particular, many LOC may simply mean that a researcher writes inefficient and repetitive code, one of the great sins of programming.
+On the contrary, a feature of good software is modularity because it enables reusing functions both inside and outside the project, resulting in fewer LOC.
 
-<!--A great example is the CRAN library that enables packages to built upon other packages and reuse their functionality.
-However, if LOC were a metric that a researcher are compared with, it encourages researchers to copy&paste chunks of code from other packages instead of importing them in their project.
-<!--Second, lines of code is an ambiguous metric.
-It can be calculated differently depending on whether comments do count or whether physical lines of code or logical lines of code are counted.-->
-
-## ID 7: Contributer Roles and Involvement
+## ID 7: Contributor Roles and Involvement
 <!-- Larger software project, but the same effort of the individual researcher leads to lower scores? -->
@@
 ## ID 8: License
-At present, the question of whether the software is open source is not used in the computation of a threshold in the prescreening phase.
-However, we think an open license is central for assessing both rigor and (potentially) increasing impact and should be part of the phase I assessment.
-Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by re-using software in other packages.
-Broader impact in science is more likely with open source software.
-In addition, many aspects of rigor are not possible to evaluate for closed-source software --- for example, whether it is well-tested or whether bugs have been fixed.
-Therefore, we propose to penalize software if it does not adhere to an open-source licenses (those approved in a review process by the Open Source Society (https://opensource.org/licenses) by allowing them only half of the achievable total points.
-<!--Alternatively, one may indirectly score openness by assessing the indicators of rigor as discussed in the second part of this commentary, where it is only possible to gain points for certain aspects of rigor if the software is open-source.-->
+At present, whether the software is open source is not used in the computation of a threshold in the prescreening phase.
+However, an open license is central for assessing both rigor and (potentially) increasing impact and should be part of the phase I assessment.
+Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by reusing software in other packages.
+Broader impact in science is more likely with open-source software.
+In addition, many aspects of rigor are not possible to evaluate for closed-source software --- for example, whether it is well-tested or bugs have been fixed.
+Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the Open Source Initiative (https://opensource.org/licenses) by allowing them only half of the total achievable points.
 
@@
 
-This is one of only two criteria used in the prescreening phase of the proposal, and therefore is of central importance.
-This item is important because it assesses aspects of rigor in software development: documentation, active maintenance, and testing.
+This is one of only two criteria used in the prescreening phase of the proposal and therefore is of central importance.
+It is furthermore important because it assesses aspects of rigor in software development: documentation, active maintenance, and testing.
 However, by incorporating the size of the user base, it confounds *usablity* (as an aspect of rigor) with *usage* (as an aspect of impact).
-Importantly, the criteria for the different proposed categories are not clearly defined.
+In addition, the criteria for the different proposed categories are not clearly defined.
 For example, the difference between "fairly extensive" and "extensive" documentation is unclear.
@@
 
-Tests are essential to discover incorrect functionality, investigate scalability of code and reveal poor design choices.
-There are a variety of useful tests, such as unit tests of subcomponents or tests of software functionality at larger scale (e.g., see the testthat package in R, @wickham2011).
-It would be possible to quantify aspects of software testing, for example, by assessing code coverage defined as the percentage of code lines that are executed during testing.
-However, we believe that we should give one point for software that promises that major functionality is covered by tests.
-<!--They are usually separated into different categories:
-*Unit tests* evaluate only small pieces of code, for example specific R-functions.
-*Functional Tests* test whether the input of a user interacting with the software leads to the correct output.
-There are many more types of tests, and the type of tests appropriate for a specific software product depends on the software itself.
-However, we argue it should at least be tested whether the software can be installed on different platforms, and whether the core functionality of the software is executed without errors (*Smoke Test*).
-In addition, it might be possible to give higher scores if different types of tests are employed, or to define a minimum of tests types that should be present, for example *unit tests*, *integration tests* and *functional tests*.
-Another metric that could be assessed is code coverage, defined as the percentage of code lines that are executed during testing.
-However, we would like to note that this metric is to be taken with a grain of salt, as the quality and number of the tests is not taken into account.
-Tests should be automated, or at least reproducible.
-This rules out closed-source software tests.
--->
+Tests are essential to discover incorrect functionality, investigate code scalability and reveal poor design choices.
+There are a variety of useful tests, such as unit tests of subcomponents or tests of software functionality at a larger scale (e.g., see the testthat package in R, @wickham2011).
+It is possible to quantify aspects of software testing, for example, by assessing code coverage, defined as the percentage of code lines executed during testing.
+However, we believe that we should give points for software that promises that major functionality is covered by tests.
+Those tests should be automated or at least open-source and reproducible.
+
 ## Documentation
@@
 Just like for tests, there are different types of documentation.
-For example, there are tutorials showcasing the usage of software with examples, and there is application programming interface (API) documentation for individual functions and classes to enable reuse of functions in other software packages.
-We propose to identify relevant categories of documentation for research software and score the presence of each of them separately.
-<!-- maybe propose the ones from joss: Installation Instructions, Tutorials, API, Community Guidelines -->
-<!--AB: moved this to maintenance; Another aspect of documentation is the possibility to ask questions to the package authors (see for example the lavaan mailing list).
--->
+For example, tutorials showcase software usage with examples, and there is application programming interface (API) documentation for individual functions and classes to enable reusing functions in other software packages.
+We propose to identify relevant categories of documentation for research software, e.g., installation instructions, tutorials, API, and community guidelines <!-- Cite JOSS!-->, and score the presence of each of them separately.
 
@@
 This should be reflected in the assessment procedure.
-We propose to score two aspects of maintenance separately, maintaining the actual code base (such as active bug fixing, documenting changes in logs), and maintaining the community (such as providing the possibility to report bugs, feature requests, or support requests via tickets or mailing lists).
+We propose to score two aspects of maintenance separately, maintaining the code base (such as active bug fixing and documenting changes in logs) and maintaining the community (such as providing the possibility to report bugs, feature requests, or support requests via tickets or mailing lists).
 
-## Measuring Impact
+# Measuring Impact
 
-In our opinion, total citation metrics, number of users, downloads per month, github stars and similar may provide a coarse measure for the impact of a software package, even though it is important to note the shortcomings we described above.
+Total citation metrics, number of users, downloads per month, GitHub stars and similar may provide a coarse measure for the impact of a software package, even though it is important to note the shortcomings we described above.
 
@@
 
-<!--Complementing the quantitative aspect of impact, we suggest to introduce a mandatory qualitative judgement on the impact of the software, which fits with what @gartner2022responsible proposed as indicator "merit / impact statement".-->
-
-In our view, researchers should be requested to indicate at least one (and up to three)  research projects that directly benefited from the software contributions they made.
-We believe this is a fair assessment of the real impact no matter whether someone contributed documentation to run certain types of models, whether someone contributed a specific feature that enabled a given research project to use the software, or whether someone is the primary author of the entire package.
+In our view, researchers should be requested to indicate at least one (and up to three)  research projects that directly benefited from their software contributions.
+We believe this to be a fair assessment of the actual impact, independent on the specific contribution the individual researcher made.
 
@@
 In sum, we are thankful to @gartner2022responsible for highlighting the importance of research software contributions as scientifically valuable products.
-We believe that the current proposal must aim to better reflect the distinction between rigor and impact.
-To this end, we emphasize a more fine-grained assessment of rigor and a qualitative statement on impact, in which software authors should argue in how far their project paved the way for other scientific endeavours.
-Last, we hope to arrive at an evaluation scheme that incentivises the development of scientific open source software.
+We believe the current proposal must aim to better reflect the distinction between rigor and impact for software, similar to what was proposed to evaluate journal articles.
+To this end, we emphasize a more fine-grained assessment of rigor and a qualitative statement on impact, in which software authors should argue in how far their project paved the way for other scientific endeavors.
+Last, we hope to arrive at an evaluation scheme that incentivizes the development of scientific open-source software.
 

### 2023-02-23 manuscript.Rmd
Update manuscript.Rmd
British English

@@
 "Software is an essential research product, and the effort to produce, maintain, adapt, and curate code must be recognized."
-However, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivize software development and, specifically, good software engineering practice [@baxter2012research].
+However, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivise software development and, specifically, good software engineering practice [@baxter2012research].
 In line with principle 3 of @schonbrodt2022responsible, we suggest that criteria for research software contributions must capture two major dimensions: rigor and impact.
@@
 To this end, we emphasize a more fine-grained assessment of rigor and a qualitative statement on impact, in which software authors should argue in how far their project paved the way for other scientific endeavors.
-Last, we hope to arrive at an evaluation scheme that incentivizes the development of scientific open-source software.
+Last, we hope to arrive at an evaluation scheme that incentivises the development of scientific open-source software.
 

### 2023-02-23 manuscript.Rmd
Update manuscript.Rmd
more BE :)

@@
 We believe the current proposal must aim to better reflect the distinction between rigor and impact for software, similar to what was proposed to evaluate journal articles.
-To this end, we emphasize a more fine-grained assessment of rigor and a qualitative statement on impact, in which software authors should argue in how far their project paved the way for other scientific endeavors.
+To this end, we emphasize a more fine-grained assessment of rigor and a qualitative statement on impact, in which software authors should argue in how far their project paved the way for other scientific endeavours.
 Last, we hope to arrive at an evaluation scheme that incentivises the development of scientific open-source software.

### 2023-02-23 manuscript.Rmd

@@
 This package has more than 58,000 citations on Google scholar, yet it is unlikely that researchers will cite the underlying optimization algorithm.
-For another example, the pdc package [@pdc-paper] offers functions to cluster time series based on relative complexity.
-<!-- Here is a sentence missing about why this is an example that shows the invalidity of citation based metrics -->
+For another example, the pdc package [@pdc-paper] offers functions to cluster time series based on one specific algorithm.
+The TSclust package [@TSclust] is a wrapper package, which imports and makes accessible functionality from the pdc package together with various other clustering approaches, which is very useful from a users' perspective; however, we noticed that researchers cite TSclust instead of pdc, which makes citation-based metrics less valid.
 

### 2023-02-24 manuscript.Rmd
Update manuscript.Rmd

@@
 ## ID 7: Contributor Roles and Involvement
-<!-- Larger software project, but the same effort of the individual researcher leads to lower scores? -->
+
+We support the standardised assessment of project contributor roles, similar to the Contributor Roles Taxonomy (CRediT; https://credit.niso.org). 
+However, we like to point out that the current evaluation schemes yields lower scores for the same effort of the individual researcher if they are part of a larger software project.
 

### 2023-03-01 manuscript.Rmd
add repro metadata
missing newline

@@
 output            : papaja::apa6_pdf
-
 equal_contrib: yes
-
 header-includes: |
   \usepackage{hyperref}
+repro:
+  packages:
+    - papaja
+    - aaronpeikert/repro
 ---
@@
 # Introduction
+
 Based on four principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a specific evaluation scheme for published manuscripts, reusable datasets, and research software.

### 2023-03-01 manuscript.Rmd
prettify here and there

@@
 Further, citations suffer from serious shortcomings when used to evaluate the impact of research software, particularly when used comparatively in the context of hire and promotion.
-For example, citing data analysis packages is much more commonly accepted than citing supporting packages (such as papaja [@R-papaja] used to render this article).
+For example, citing data analysis packages is much more commonly accepted than citing supporting packages [such as papaja, @R-papaja, used to render this article].
 Further, functionality of successful, modular scientific software is ideally reused in other software packages to avoid code dublication and enable faster development of new software.
@@
 NLopt is a backbone for many scientific packages both because it implements various optimization algorithms but also because it is open-source and can inspire re-implementations of these algorithms.
-For example, consider the famous lme4 package [@bates2014fitting] for generalized linear mixed-effects models, that is partly based on NLopt.
+One such example is the famous lme4 package [@bates2014fitting] for generalized linear mixed-effects models, that is partly based on NLopt.
 This package has more than 58,000 citations on Google scholar, yet it is unlikely that researchers will cite the underlying optimization algorithm.
@@
 ## ID 8: License
+
 At present, whether the software is open source is not used in the computation of a threshold in the prescreening phase.
@@
 Tests are essential to discover incorrect functionality, investigate code scalability and reveal poor design choices.
-There are a variety of useful tests, such as unit tests of subcomponents or tests of software functionality at a larger scale (e.g., see the testthat package in R, @wickham2011).
+There are a variety of useful tests, such as unit tests of subcomponents or tests of software functionality at a larger scale [e.g., see the testthat package in R, @wickham2011].
 It is possible to quantify aspects of software testing, for example, by assessing code coverage, defined as the percentage of code lines executed during testing.
@@
 We believe that the suggested merit statement is most useful to assess impact of research software and that this should be the primary statement for committee members to evaluate if they are less concerned with the technical aspects of research software development.
-
 In our view, researchers should be requested to indicate at least one (and up to three)  research projects that directly benefited from their software contributions.

### 2023-03-01 manuscript.Rmd
merge two sections to begin more impactfull and reduce wordcount

--- 2023-03-01 manuscript.Rmd
@@
 
-## ID 5: Date of first fully functional public version
+## ID 5: Date of first fully functional public version/ID 9: Citations
 
-The date of the first fully functional public version serves as a denominator in computing the number of citations (see ID 9) relative to software age.
-In our experience, research software often evolves driven by time constraints and the need for specific functionality, not according to feature completeness. 
-Thus, it may generally suffice to ask for the age of the software relative to its first public release without any constraints on functionality.
+Computing the citations in relation to age of software seems to be inconsistent with the proposal of @schonbrodt2022responsible, who reminded us that a core principle of the implementation of DORA is "abandon the use of invalid quantitative metrics of research quality and productivity in hiring and promotion" (p.2).
+It is unclear why an ad-hoc created quantitative metric for software would be in any way more valid than the equivalent for articles.
+In fact, there are several additional shortcomings in relation to software.
+We discuss shortcomings of the numerator (number of citations) here and the issues of the denominator (software age) in the next section.
 
-## ID 6: Date of most recent substantive update
-
-We believe it is not always clear when updates are considered 'substantive.'
-As an important aspect of rigor, software needs to be actively maintained.
-To assess this, we propose to provide a checkbox in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
-In addition, authors have a chance to explain why their active maintenance may be different from these guidelines.
-
-## ID 9: Citations
-
-@gartner2022responsible suggest the number of citations to evaluate the impact of a software package.
-First, we like to point out an inconsistency in proposing citation-based metrics and the propositions of @schonbrodt2022responsible, who reminded us that a core principle of the implementation of DORA is "abandon the use of invalid quantitative metrics of research quality and productivity in hiring and promotion" (p.2) and instead to focus on individual achievements.
-Further, citations suffer from serious shortcomings when used to evaluate the impact of research software, particularly when used comparatively in the context of hire and promotion.
+Citations suffer from serious shortcomings when used to evaluate the impact of research software, particularly when used comparatively in the context of hire and promotion.
 For example, citing data analysis packages is much more commonly accepted than citing supporting packages (such as papaja [@R-papaja] used to render this article).
@@
 The TSclust package [@TSclust] is a wrapper package, which imports and makes accessible functionality from the pdc package together with various other clustering approaches, which is very useful from a users' perspective; however, we noticed that researchers cite TSclust instead of pdc, which makes citation-based metrics less valid.
+
+## ID 6: Date of most recent substantive update
+
+Both ID 5 and 6 are difficult to ascertain because it is not always clear when updates are considered 'substantive.' or software 'fully functional'.
+We appreciate the importance of assessing maintenance as part of rigor.
+To assess this, we propose to provide a checkbox in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
+In addition, authors have a chance to explain why their active maintenance may be different from these guidelines.
 

### 2023-03-01 manuscript.Rmd
add required gh packages
forgot here package
pretty again

--- 2023-03-01 manuscript.Rmd
@@
 While we highly encourage this from both a software engineering perspective and for scientific progress, it challenges the validity of citation-based metrics for impact.
-For example, consider the NLopt optimization suite, which counts 1,711 citations on Google scholar at the time of writing.
+For example, consider the `NLopt` optimization suite, which counts 1,711 citations on Google scholar at the time of writing.
 NLopt is a backbone for many scientific packages both because it implements various optimization algorithms but also because it is open-source and can inspire re-implementations of these algorithms.
-One such example is the famous lme4 package [@bates2014fitting] for generalized linear mixed-effects models, that is partly based on NLopt.
-This package has more than 58,000 citations on Google scholar, yet it is unlikely that researchers will cite the underlying optimization algorithm.
+One such example is the famous `lme4` package [@bates2014fitting] for generalized linear mixed-effects models, that is partly based on `NLopt`.
+`lme4` has more than 58,000 citations on Google scholar, yet it is unlikely that researchers will cite the underlying optimization algorithm.
 For another example, the pdc package [@pdc-paper] offers functions to cluster time series based on one specific algorithm.

### 2023-03-02 manuscript.Rmd

--- 2023-03-01 manuscript.Rmd
@@
     - papaja
-    - aaronpeikert/repro
+    - here
+    - aaronpeikert/repro@5075336
+    - benmarwick/wordcountaddin@e075baa
 ---
@@
 
-Citations suffer from serious shortcomings when used to evaluate the impact of research software, particularly when used comparatively in the context of hire and promotion.
-For example, citing data analysis packages is much more commonly accepted than citing supporting packages (such as papaja [@R-papaja] used to render this article).
+Citations suffer from serious shortcomings when used to evaluate the impact of research software, particularly when used comparatively in the context of hiring and promotion.
+For example, citing data analysis packages is much more commonly accepted than citing supporting packages [such as papaja, @R-papaja, used to render this article].
+
 Further, functionality of successful, modular scientific software is ideally reused in other software packages to avoid code dublication and enable faster development of new software.
 While we highly encourage this from both a software engineering perspective and for scientific progress, it challenges the validity of citation-based metrics for impact.
-For example, consider the NLopt optimization suite, which counts 1,711 citations on Google scholar at the time of writing.
+For example, consider the `NLopt` optimization suite, which counts 1,711 citations on Google scholar at the time of writing.
 NLopt is a backbone for many scientific packages both because it implements various optimization algorithms but also because it is open-source and can inspire re-implementations of these algorithms.
-For example, consider the famous lme4 package [@bates2014fitting] for generalized linear mixed-effects models, that is partly based on NLopt.
-This package has more than 58,000 citations on Google scholar, yet it is unlikely that researchers will cite the underlying optimization algorithm.
+One such example is the famous `lme4` package [@bates2014fitting] for generalized linear mixed-effects models, that is partly based on `NLopt`.
+`lme4` has more than 58,000 citations on Google scholar, yet it is unlikely that researchers will cite the underlying optimization algorithm.
 For another example, the pdc package [@pdc-paper] offers functions to cluster time series based on one specific algorithm.

### 2023-03-02 manuscript.Rmd
added repro link
minor fix

@@
 
-At present, whether the software is open source is not used in the computation of a threshold in the prescreening phase.
-However, an open license is central for assessing both rigor and (potentially) increasing impact and should be part of the phase I assessment.
+At present, whether a piece of software is open source is not evaluated in the prescreening phase.
+However, an open license is central for assessing both rigor and (potential) impact, and should be part of the phase I assessment.
 Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by reusing software in other packages.
-Broader impact in science is more likely with open-source software.
-In addition, many aspects of rigor are not possible to evaluate for closed-source software --- for example, whether it is well-tested or bugs have been fixed.
+Open-source software makes broader impact more likely and is a prerequisite for full transparency and reproducibility [@peikert2021reproducible].
+In addition, many aspects of rigor are impossible to evaluate for closed-source software --- for example, whether it is well-tested or bugs have been fixed.
 Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the Open Source Initiative (https://opensource.org/licenses) by allowing them only half of the total achievable points.

### 2023-03-02 manuscript.Rmd
moved manifesto citation to URL citation

@@
 
-Based on four principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a specific evaluation scheme for published manuscripts, reusable datasets, and research software.
-We specifically endorse the increased emphasis on research software as a recognizable and laudable scientific contribution.
+Based on four principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a concrete evaluation scheme for published manuscripts, reusable datasets, and research software.
+We strongly support the increased emphasis on research software as a creditable and commendable scientific contribution.
 Why are research software contributions important scientific contributions?
-We would like to respond with a quote from the Science Code Manifesto (http://sciencecodemanifesto.org/):
+We would like to respond with a quote from the Science Code Manifesto [@manifesto]:
 "Software is a cornerstone of science. Without software, twenty-first century science would be impossible."

### 2023-03-02 manuscript.Rmd
rigour (BE) also some minor changes

@@
 
-keywords          : "research software, open science, metrics, rigor, impact"
+keywords          : "research software, open science, metrics, rigour, impact"
 wordcount         : "`r tryCatch(wordcountaddin::word_count(here::here('manuscript.Rmd')))`"
@@
 However, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivise software development and, specifically, good software engineering practice [@baxter2012research].
-In line with principle 3 of @schonbrodt2022responsible, we suggest that criteria for research software contributions must capture two major dimensions: rigor and impact.
+In line with principle 3 of @schonbrodt2022responsible, we suggest that criteria for research software contributions must capture two major dimensions: rigour and impact.
 Impact measures whether the scholarly effort, implementation, and dissemination actually had a visible effect on the field.
-Rigor means implementing high standards and best practices for ensuring transparency, correctness, and reusability of a piece of software.
-By setting a high bar of rigor in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and, thus, better science.
-From this perspective, we comment on the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible) and propose alternatives to better capture rigor and impact.
+rigour means implementing high standards and best practices for ensuring transparency, correctness, and reusability of a piece of software.
+By setting a high bar of rigour in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and, thus, better science.
+From this perspective, we comment on some indicators of the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible).
 
@@
 
-Computing the citations in relation to age of software seems to be inconsistent with the proposal of @schonbrodt2022responsible, who reminded us that a core principle of the implementation of DORA is "abandon the use of invalid quantitative metrics of research quality and productivity in hiring and promotion" (p.2).
-It is unclear why an ad-hoc created quantitative metric for software would be in any way more valid than the equivalent for articles.
+Computing the citations in relation to age of software seems to be inconsistent with the proposal of @schonbrodt2022responsible, who reminded us that a core principle of the implementation of DORA is to "abandon the use of invalid quantitative metrics of research quality and productivity in hiring and promotion" (p.2).
+It is unclear why a citation-based metric for software would be more valid than the equivalent for articles.
 In fact, there are several additional shortcomings in relation to software.
@@
 
-Further, functionality of successful, modular scientific software is ideally reused in other software packages to avoid code dublication and enable faster development of new software.
-While we highly encourage this from both a software engineering perspective and for scientific progress, it challenges the validity of citation-based metrics for impact.
+Further, functionality of successful, modular scientific software is ideally reused in other software packages to avoid code duplication and enable faster development of new software.
+While we highly encourage reuse from both a software engineering perspective and for scientific progress, it challenges the validity of citation-based metrics for impact.
 For example, consider the `NLopt` optimization suite, which counts 1,711 citations on Google scholar at the time of writing.
-NLopt is a backbone for many scientific packages both because it implements various optimization algorithms but also because it is open-source and can inspire re-implementations of these algorithms.
+`NLopt` is a backbone for many scientific packages both because it implements various optimization algorithms but also because it is open-source and can inspire re-implementations of these algorithms.
 One such example is the famous `lme4` package [@bates2014fitting] for generalized linear mixed-effects models, that is partly based on `NLopt`.
 `lme4` has more than 58,000 citations on Google scholar, yet it is unlikely that researchers will cite the underlying optimization algorithm.
-For another example, the pdc package [@pdc-paper] offers functions to cluster time series based on one specific algorithm.
-The TSclust package [@TSclust] is a wrapper package, which imports and makes accessible functionality from the pdc package together with various other clustering approaches, which is very useful from a users' perspective; however, we noticed that researchers cite TSclust instead of pdc, which makes citation-based metrics less valid.
+For another example, the `pdc` package [@pdc-paper] offers functions to cluster time series based on one specific algorithm.
+The `TSclust` package [@TSclust] is a wrapper package, which imports and makes accessible functionality from the `pdc` package as well as various other clustering approaches, which is very useful from a users' perspective; however, we noticed that researchers now cite `TSclust` instead of `pdc`, which challenges the citation-based assessment of impact.
 
@@
 
-Both ID 5 and 6 are difficult to ascertain because it is not always clear when updates are considered 'substantive.' or software 'fully functional'.
-We appreciate the importance of assessing maintenance as part of rigor.
-To assess this, we propose to provide a checkbox in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
+Both ID 5 and 6 are difficult to ascertain because it is not always clear when updates are considered 'substantive' or software 'fully functional'.
+We appreciate the importance of assessing maintenance as part of rigour.
+To assess this, we propose to provide a simple checkbox, in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
 In addition, authors have a chance to explain why their active maintenance may be different from these guidelines.
@@
 At present, whether a piece of software is open source is not evaluated in the prescreening phase.
-However, an open license is central for assessing both rigor and (potential) impact, and should be part of the phase I assessment.
+However, an open license is central for assessing both rigour and (potential) impact, and should be part of the phase I assessment.
 Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by reusing software in other packages.
 Open-source software makes broader impact more likely and is a prerequisite for full transparency and reproducibility [@peikert2021reproducible].
-In addition, many aspects of rigor are impossible to evaluate for closed-source software --- for example, whether it is well-tested or bugs have been fixed.
+In addition, many aspects of rigour are impossible to evaluate for closed-source software --- for example, whether it is well-tested or bugs have been fixed.
 Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the Open Source Initiative (https://opensource.org/licenses) by allowing them only half of the total achievable points.
@@
 This is one of only two criteria used in the prescreening phase of the proposal and therefore is of central importance.
-It is furthermore important because it assesses aspects of rigor in software development: documentation, active maintenance, and testing.
-However, by incorporating the size of the user base, it confounds *usablity* (as an aspect of rigor) with *usage* (as an aspect of impact).
+It is also important because it assesses aspects of rigour in software development: documentation, active maintenance, and testing.
+However, by incorporating the size of the user base, it confounds *usablity* (as an aspect of rigour) with *usage* (as an aspect of impact).
 In addition, the criteria for the different proposed categories are not clearly defined.
@@
 As a result, this indicator is more of a "gut-feeling" indicator, roughly assessing the "size" of the software project.
-Instead, in the following we propose to assess rigor and impact independently as primary aspects of a software contribution.
+Instead, in the following we propose to assess rigour and impact independently as primary aspects of a software contribution.
 
-# Rigor for software implementations
+# rigour for software implementations
 
 <!--In the following, we discuss aspects that we believe should be scored in phase one of the assessment.-->
-We propose to use the following aspects as equally weighted indicators of rigor in lieu of the proposed coarse reusability indicator (item 17 in Table 3 of the proposal).
+We propose to use the following aspects as equally weighted indicators of rigour instead of the proposed broader reusability indicator (item 17 in Table 3 of the proposal).
 <!-- some introductory text missing -->
@@
 In our view, researchers should be requested to indicate at least one (and up to three)  research projects that directly benefited from their software contributions.
-We believe this to be a fair assessment of the actual impact, independent on the specific contribution the individual researcher made.
+We believe this to be a fair assessment of the actual impact the specific contribution of the individual researcher had.
 
@@
 In sum, we are thankful to @gartner2022responsible for highlighting the importance of research software contributions as scientifically valuable products.
-We believe the current proposal must aim to better reflect the distinction between rigor and impact for software, similar to what was proposed to evaluate journal articles.
-To this end, we emphasize a more fine-grained assessment of rigor and a qualitative statement on impact, in which software authors should argue in how far their project paved the way for other scientific endeavours.
+We believe the current proposal should aim to better reflect the distinction between rigour and impact for software, similar to the guidelines proposed to evaluate journal articles.
+To this end, we suggest a more fine-grained assessment of rigour, and put emphasis on the merit statement, in which software authors should argue in how far their project impacted other scientific endeavours.
 Last, we hope to arrive at an evaluation scheme that incentivises the development of scientific open-source software.

### 2023-03-02 manuscript.Rmd
added JOSS link BE -> AE

@@
 
-keywords          : "research software, open science, metrics, rigour, impact"
+keywords          : "research software, open science, metrics, rigor, impact"
 wordcount         : "`r tryCatch(wordcountaddin::word_count(here::here('manuscript.Rmd')))`"
@@
 
-Based on four principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a concrete evaluation scheme for published manuscripts, reusable datasets, and research software.
+Based on four principles of a more responsible research assessment in academic hiring and promotion processes [@schonbrodt2022responsible], @gartner2022responsible suggested a concrete evaluation scheme for published manuscripts, reusable data sets, and research software.
 We strongly support the increased emphasis on research software as a creditable and commendable scientific contribution.
@@
 "Software is an essential research product, and the effort to produce, maintain, adapt, and curate code must be recognized."
-However, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivise software development and, specifically, good software engineering practice [@baxter2012research].
-In line with principle 3 of @schonbrodt2022responsible, we suggest that criteria for research software contributions must capture two major dimensions: rigour and impact.
+However, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivize software development and, specifically, good software engineering practice [@baxter2012research].
+In line with principle 3 of @schonbrodt2022responsible, we suggest that criteria for research software contributions must capture two major dimensions: rigor and impact.
 Impact measures whether the scholarly effort, implementation, and dissemination actually had a visible effect on the field.
-rigour means implementing high standards and best practices for ensuring transparency, correctness, and reusability of a piece of software.
-By setting a high bar of rigour in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and, thus, better science.
+rigor means implementing high standards and best practices for ensuring transparency, correctness, and re-usability of a piece of software.
+By setting a high bar of rigor in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and, thus, better science.
 From this perspective, we comment on some indicators of the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible).
@@
 Both ID 5 and 6 are difficult to ascertain because it is not always clear when updates are considered 'substantive' or software 'fully functional'.
-We appreciate the importance of assessing maintenance as part of rigour.
-To assess this, we propose to provide a simple checkbox, in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
+We appreciate the importance of assessing maintenance as part of rigor.
+To assess this, we propose to provide a simple check box, in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
 In addition, authors have a chance to explain why their active maintenance may be different from these guidelines.
@@
 
-We support the standardised assessment of project contributor roles, similar to the Contributor Roles Taxonomy (CRediT; https://credit.niso.org). 
+We support the standardized assessment of project contributor roles, similar to the Contributor Roles Taxonomy (CRediT; https://credit.niso.org). 
 However, we like to point out that the current evaluation schemes yields lower scores for the same effort of the individual researcher if they are part of a larger software project.
@@
 
-At present, whether a piece of software is open source is not evaluated in the prescreening phase.
-However, an open license is central for assessing both rigour and (potential) impact, and should be part of the phase I assessment.
+At present, whether a piece of software is open source is not evaluated in the pre-screening phase.
+However, an open license is central for assessing both rigor and (potential) impact, and should be part of the phase I assessment.
 Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by reusing software in other packages.
 Open-source software makes broader impact more likely and is a prerequisite for full transparency and reproducibility [@peikert2021reproducible].
-In addition, many aspects of rigour are impossible to evaluate for closed-source software --- for example, whether it is well-tested or bugs have been fixed.
+In addition, many aspects of rigor are impossible to evaluate for closed-source software --- for example, whether it is well-tested or bugs have been fixed.
 Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the Open Source Initiative (https://opensource.org/licenses) by allowing them only half of the total achievable points.
@@
 
-This is one of only two criteria used in the prescreening phase of the proposal and therefore is of central importance.
-It is also important because it assesses aspects of rigour in software development: documentation, active maintenance, and testing.
-However, by incorporating the size of the user base, it confounds *usablity* (as an aspect of rigour) with *usage* (as an aspect of impact).
+This is one of only two criteria used in the pre-screening phase of the proposal and therefore is of central importance.
+It is also important because it assesses aspects of rigor in software development: documentation, active maintenance, and testing.
+However, by incorporating the size of the user base, it confounds *usability* (as an aspect of rigor) with *usage* (as an aspect of impact).
 In addition, the criteria for the different proposed categories are not clearly defined.
@@
 As a result, this indicator is more of a "gut-feeling" indicator, roughly assessing the "size" of the software project.
-Instead, in the following we propose to assess rigour and impact independently as primary aspects of a software contribution.
+Instead, in the following we propose to assess rigor and impact independently as primary aspects of a software contribution.
 
-# rigour for software implementations
+# rigor for software implementations
 
 <!--In the following, we discuss aspects that we believe should be scored in phase one of the assessment.-->
-We propose to use the following aspects as equally weighted indicators of rigour instead of the proposed broader reusability indicator (item 17 in Table 3 of the proposal).
+We propose to use the following aspects as equally weighted indicators of rigor instead of the proposed broader re-usability indicator (item 17 in Table 3 of the proposal).
 <!-- some introductory text missing -->
@@
 Tests are essential to discover incorrect functionality, investigate code scalability and reveal poor design choices.
-There are a variety of useful tests, such as unit tests of subcomponents or tests of software functionality at a larger scale [e.g., see the testthat package in R, @wickham2011].
+There are a variety of useful tests, such as unit tests of sub components or tests of software functionality at a larger scale [e.g., see the testthat package in R, @wickham2011].
 It is possible to quantify aspects of software testing, for example, by assessing code coverage, defined as the percentage of code lines executed during testing.
@@
 For example, tutorials showcase software usage with examples, and there is application programming interface (API) documentation for individual functions and classes to enable reusing functions in other software packages.
-We propose to identify relevant categories of documentation for research software, e.g., installation instructions, tutorials, API, and community guidelines <!-- Cite JOSS!-->, and score the presence of each of them separately.
+We propose to identify relevant categories of documentation for research software, e.g., installation instructions, tutorials, API, and community guidelines (also see https://joss.readthedocs.io/en/latest/review_checklist.html), and score the presence of each of them separately.
 
@@
 In sum, we are thankful to @gartner2022responsible for highlighting the importance of research software contributions as scientifically valuable products.
-We believe the current proposal should aim to better reflect the distinction between rigour and impact for software, similar to the guidelines proposed to evaluate journal articles.
-To this end, we suggest a more fine-grained assessment of rigour, and put emphasis on the merit statement, in which software authors should argue in how far their project impacted other scientific endeavours.
-Last, we hope to arrive at an evaluation scheme that incentivises the development of scientific open-source software.
+We believe the current proposal should aim to better reflect the distinction between rigor and impact for software, similar to the guidelines proposed to evaluate journal articles.
+To this end, we suggest a more fine-grained assessment of rigor, and put emphasis on the merit statement, in which software authors should argue in how far their project impacted other scientific endeavors.
+Last, we hope to arrive at an evaluation scheme that incentivizes the development of scientific open-source software.
 

### 2023-03-02 manuscript.Rmd
added submission note to front page as requested by journal

@@
 
-  To be submitted to: Meta-Psychology
+  Submitted to Meta-Psychology.
+Participate in open peer review by commenting through hypothes.is directly on this preprint.
+The full editorial process of all articles under review at Meta-Psychology can be found following this link: https://tinyurl.com/mp-submissions.
+You will find this preprint by searching for the first author’sname.
 

### 2023-03-02 manuscript.Rmd
last changes

--- 2023-03-02 manuscript.Rmd
@@
 Impact measures whether the scholarly effort, implementation, and dissemination actually had a visible effect on the field.
-rigour means implementing high standards and best practices for ensuring transparency, correctness, and reusability of a piece of software.
+Rigour means implementing high standards and best practices for ensuring transparency, correctness, and reusability of a piece of software.
 By setting a high bar of rigour in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and, thus, better science.
@@
 
-## ID 5: Date of first fully functional public version/ID 9: Citations
+## ID 9: Citations/ID 5: Date of first fully functional public version
 
@@
 We appreciate the importance of assessing maintenance as part of rigour.
+However, the current item asks for updates that add "substantive functionality", while maintenance work (e.g., bug fixes) corresponds to patch releases.
 To assess this, we propose to provide a simple checkbox, in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
@@
 In addition, many aspects of rigour are impossible to evaluate for closed-source software --- for example, whether it is well-tested or bugs have been fixed.
-Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the Open Source Initiative (https://opensource.org/licenses) by allowing them only half of the total achievable points.
+Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the Open Source Initiative (https://opensource.org/licenses)) by allowing them only half of the total achievable points.
 
@@
 
-# rigour for software implementations
+# Rigour for Research Software
 

### 2023-03-03 manuscript.Rmd

--- 2023-03-02 manuscript.Rmd
@@
 Impact measures whether the scholarly effort, implementation, and dissemination actually had a visible effect on the field.
-rigor means implementing high standards and best practices for ensuring transparency, correctness, and re-usability of a piece of software.
+Rigor means implementing high standards and best practices for ensuring transparency, correctness, and re-usability of a piece of software.
 By setting a high bar of rigor in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and, thus, better science.
@@
 Both ID 5 and 6 are difficult to ascertain because it is not always clear when updates are considered 'substantive' or software 'fully functional'.
-We appreciate the importance of assessing maintenance as part of rigor.
-To assess this, we propose to provide a simple check box, in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
+To assess active maintenance as an aspect of rigor, we propose to provide a simple checkbox, in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
 In addition, authors have a chance to explain why their active maintenance may be different from these guidelines.
@@
 
-# rigor for software implementations
+# Assessing Rigor
 

### 2023-03-03 manuscript.Rmd
changed meta-psychology header
Update manuscript.Rmd

@@
   Submitted to Meta-Psychology.
+
   Participate in open peer review by commenting through hypothes.is directly on this preprint.
 The full editorial process of all articles under review at Meta-Psychology can be found following this link: https://tinyurl.com/mp-submissions.
-You will find this preprint by searching for the first author’sname.
+You will find this preprint by searching for the first author’s name.
 
@@
 We would like to respond with a quote from the Science Code Manifesto [@manifesto]:
-"Software is a cornerstone of science. Without software, twenty-first century science would be impossible."
+"Software is a cornerstone of science. Without software, twenty-first century science would be impossible"
 and
-"Software is an essential research product, and the effort to produce, maintain, adapt, and curate code must be recognized."
+"software is an essential research product, and the effort to produce, maintain, adapt, and curate code must be recognized."
 However, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivize software development and, specifically, good software engineering practice [@baxter2012research].

### 2023-03-03 manuscript.Rmd
- important change on size/effort comment - some shortening

@@
 By setting a high bar of rigor in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and, thus, better science.
-From this perspective, we comment on some indicators of the proposed evaluation scheme for research software contributions (Table 3 of @gartner2022responsible).
+From this perspective, we comment on some indicators of the proposed evaluation scheme for research software contributions [Table 3 of @gartner2022responsible].
 
@@
 
-## ID 9: Citations/ID 5: Date of first fully functional public version
+## ID 9: Citations and ID 5: Date of first fully functional public version
 
-Computing the citations in relation to age of software seems to be inconsistent with the proposal of @schonbrodt2022responsible, who reminded us that a core principle of the implementation of DORA is to "abandon the use of invalid quantitative metrics of research quality and productivity in hiring and promotion" (p.2).
+Considering citations in relation to age of software seems to be inconsistent with the proposal of @schonbrodt2022responsible, who reminded us that a core principle of the implementation of DORA is to "abandon the use of invalid quantitative metrics of research quality and productivity in hiring and promotion" (p.2).
 It is unclear why a citation-based metric for software would be more valid than the equivalent for articles.
-In fact, there are several additional shortcomings in relation to software.
-We discuss shortcomings of the numerator (number of citations) here and the issues of the denominator (software age) in the next section.
-
-Citations suffer from serious shortcomings when used to evaluate the impact of research software, particularly when used comparatively in the context of hiring and promotion.
+In fact, citations suffer from additional shortcomings when used to evaluate the impact of research software, particularly when used comparatively in the context of hiring and promotion.
 For example, citing data analysis packages is much more commonly accepted than citing supporting packages [such as papaja, @R-papaja, used to render this article].
-
 Further, functionality of successful, modular scientific software is ideally reused in other software packages to avoid code duplication and enable faster development of new software.
@@
 `NLopt` is a backbone for many scientific packages both because it implements various optimization algorithms but also because it is open-source and can inspire re-implementations of these algorithms.
-One such example is the famous `lme4` package [@bates2014fitting] for generalized linear mixed-effects models, that is partly based on `NLopt`.
+One such example is the famous `lme4` package [@bates2014fitting] for generalized linear mixed-effects models, which is partly based on `NLopt`.
 `lme4` has more than 58,000 citations on Google scholar, yet it is unlikely that researchers will cite the underlying optimization algorithm.
 For another example, the `pdc` package [@pdc-paper] offers functions to cluster time series based on one specific algorithm.
-The `TSclust` package [@TSclust] is a wrapper package, which imports and makes accessible functionality from the `pdc` package as well as various other clustering approaches, which is very useful from a users' perspective; however, we noticed that researchers now cite `TSclust` instead of `pdc`, which challenges the citation-based assessment of impact.
+The `TSclust` package [@TSclust] is a wrapper package, which imports and makes accessible functionality from the `pdc` package as well as various other clustering approaches, which is very useful from a users' perspective; however, we noticed that researchers now cite `TSclust` instead of `pdc`.
 
@@
 Both ID 5 and 6 are difficult to ascertain because it is not always clear when updates are considered 'substantive' or software 'fully functional'.
-To assess active maintenance as an aspect of rigor, we propose to provide a simple checkbox, in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
+To assess active maintenance as an aspect of rigor, we propose to provide a simple check box, in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
 In addition, authors have a chance to explain why their active maintenance may be different from these guidelines.
@@
 We support the standardized assessment of project contributor roles, similar to the Contributor Roles Taxonomy (CRediT; https://credit.niso.org). 
-However, we like to point out that the current evaluation schemes yields lower scores for the same effort of the individual researcher if they are part of a larger software project.
+However, we like to point out that the current evaluation schemes yields different scores for the same effort of the individual researcher depending on the size of the software project.
 

### 2023-03-03 manuscript.Rmd
Update manuscript.Rmd
testthat formatting

@@
 Tests are essential to discover incorrect functionality, investigate code scalability and reveal poor design choices.
-There are a variety of useful tests, such as unit tests of sub components or tests of software functionality at a larger scale [e.g., see the testthat package in R, @wickham2011].
+There are a variety of useful tests, such as unit tests of sub components or tests of software functionality at a larger scale [e.g., see the `testthat` package in R, @wickham2011].
 It is possible to quantify aspects of software testing, for example, by assessing code coverage, defined as the percentage of code lines executed during testing.

### 2023-06-05 manuscript.Rmd
revision based on peer review comments

--- 2023-03-03 manuscript.Rmd
@@
 ---
-title             : "Commentary: 'Responsible Research Assessment II: A specific proposal for hiring and promotion in psychology'"
-shorttitle        : "Responsible research assessment"
+title             : "Assessing rigor and impact in research software"
+#subtitle          : "Commentary: 'Responsible Research Assessment II: A specific proposal for hiring and promotion in psychology'"
+shorttitle        : "Rigor and impact"
 
@@
 
-At present, whether a piece of software is open source is not evaluated in the pre-screening phase.
+At present, whether a piece of software is open source is not evaluated at the pre-screening stage.
 However, an open license is central for assessing both rigor and (potential) impact, and should be part of the phase I assessment.
@@
 In addition, many aspects of rigor are impossible to evaluate for closed-source software --- for example, whether it is well-tested or bugs have been fixed.
-Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the Open Source Initiative (https://opensource.org/licenses) by allowing them only half of the total achievable points.
+Giving more weight to open licenses aligns well with the original article's emphasis on open science.
+Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the Open Source Initiative (https://opensource.org/licenses), for example, by allowing them only half of the total achievable points.
 
@@
 In our view, researchers should be requested to indicate at least one (and up to three) research projects that directly benefited from their software contributions.
+In case of recently published research software, the merit statement can be used alternatively to describe expected benefits to the field or potential use-cases of the software.
 We believe this to be a fair assessment of the actual impact the specific contribution of the individual researcher had.
@@
 Last, we hope to arrive at an evaluation scheme that incentivizes the development of scientific open-source software.
+Note that the requirements of a given academic position should guide the weighting of research software when evaluating and selecting candidates.
+In some cases, research software development may be central to a position and therefore a requirement (such as when hiring professors of statistics or research methods), but in other cases, we hope that research software development activities at least are considered as research output comparable to published papers.
 

### 2023-06-06 manuscript.Rmd
removed whitespace
Update manuscript.Rmd
Co-authored-by: Aaron Peikert <aaron.peikert@posteo.de>

@@
 ---
-title             : "Assessing rigor and impact in research software"
+title             : "Assessing rigor and impact of research software for hiring and promotion in psychology"
 #subtitle          : "Commentary: 'Responsible Research Assessment II: A specific proposal for hiring and promotion in psychology'"

### 2023-06-06 manuscript.Rmd
Update manuscript.Rmd
Co-authored-by: Aaron Peikert <aaron.peikert@posteo.de>

--- 2023-06-05 manuscript.Rmd
@@
 ---
-title             : "Assessing rigor and impact in research software"
+title             : "Assessing rigor and impact of research software for hiring and promotion in psychology"
 #subtitle          : "Commentary: 'Responsible Research Assessment II: A specific proposal for hiring and promotion in psychology'"
-shorttitle        : "Rigor and impact"
+shorttitle        : "Assessing research software"
 

### 2023-06-06 manuscript.Rmd
Update manuscript.Rmd
Co-authored-by: Aaron Peikert <aaron.peikert@posteo.de>

@@
 Giving more weight to open licenses aligns well with the original article's emphasis on open science.
-Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the Open Source Initiative (https://opensource.org/licenses), for example, by allowing them only half of the total achievable points.
+Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the [Open Source Initiative](https://opensource.org/licenses)), for example, by allowing them only half of the total achievable points.
 

### 2023-06-06 manuscript.Rmd
Update manuscript.Rmd
Co-authored-by: Aaron Peikert <aaron.peikert@posteo.de>

@@
 Note that the requirements of a given academic position should guide the weighting of research software when evaluating and selecting candidates.
-In some cases, research software development may be central to a position and therefore a requirement (such as when hiring professors of statistics or research methods), but in other cases, we hope that research software development activities at least are considered as research output comparable to published papers.
+In some cases, research software development may be central to a position and, therefore, a requirement (such as when hiring professors of statistics or research methods), but in other cases, we hope that research software development activities at least are considered as research output comparable to published papers.
 

### 2023-06-07 manuscript.Rmd
changed abstract

--- 2023-06-06 manuscript.Rmd
@@
 ---
-title             : "Assessing rigor and impact of research software for hiring and promotion in psychology"
+title             : "Assessing rigor and impact in research software"
 #subtitle          : "Commentary: 'Responsible Research Assessment II: A specific proposal for hiring and promotion in psychology'"
-shorttitle        : "Assessing research software"
+shorttitle        : "Rigor and impact"
 
@@
 
-#abstract: |
-#  TBD
+abstract: Based on four principles of a more responsible research assessment in academic hiring and promotion processes, @gartner2022responsible suggested an evaluation scheme for published manuscripts, reusable data sets, and research software.
+This commentary responds to the proposed indicators for the evaluation of research software contributions in academic hiring and promotion processes.
+Acknowledging the significance of research software as a critical component of modern science, we propose that an evaluation scheme must emphasize the two major dimensions of rigor and impact.
+Generally, we believe that research software should be recognized as valuable scientific output in academic hiring and promotion, with the hope that this incentivizes the development of more open and better research software.
 

### 2023-06-08 manuscript.Rmd
- added "A comment on..." to title (in line with other comments) - added departments and location to affiliations

@@
 ---
-title             : "Assessing rigor and impact of research software for hiring and promotion in psychology"
+title             : "Assessing rigor and impact of research software for hiring and promotion in psychology: A comment on Gärtner et al. (2022)"
 #subtitle          : "Commentary: 'Responsible Research Assessment II: A specific proposal for hiring and promotion in psychology'"
@@
   - id            : "1"
-    institution   : "MSB Medical School Berlin"
+    institution   : "Department of Psychology, MSB Medical School Berlin, Berlin, Germany"
   - id            : "2"
-    institution   : "Max Planck Institute for Human Development"
+    institution   : "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
   - id            : "3"
-    institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research"
+    institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany"
   - id            : "4"
-    institution   : "Humboldt-Universität zu Berlin"
+    institution   : "Department of Psychology, Humboldt-Universität zu Berlin, Berlin, Germany"
 

### 2023-09-29 manuscript.tex
Deploying to gh-pages from @ brandmaier/research_software_evaluation@fc255d3fe1b2c129c478b2e3057562bfc6daab09 🚀

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{research software, open science, metrics, rigor, impact\newline\indent Word count: 1685}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}
\usepackage{hyperref}

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Assessing rigor and impact of research software for hiring and promotion in psychology: A comment on Gärtner et al.~(2022)},
  pdfauthor={Andreas M. Brandmaier1,2,3, Maximilian Ernst2,4, \& Aaron Peikert2,3,4},
  pdflang={en-EN},
  pdfkeywords={research software, open science, metrics, rigor, impact},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Assessing rigor and impact of research software for hiring and promotion in psychology: A comment on Gärtner et al.~(2022)}
\author{Andreas M. Brandmaier\textsuperscript{1,2,3}, Maximilian Ernst\textsuperscript{2,4}, \& Aaron Peikert\textsuperscript{2,3,4}}
\date{}

\shorttitle{Assessing research software}

\authornote{

The authors made the following contributions. Andreas M. Brandmaier: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing, Supervision; Maximilian Ernst: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing; Aaron Peikert: Conceptualization, Writing - Original Draft Preparation, Writing - Review \& Editing.

Correspondence concerning this article should be addressed to Andreas M. Brandmaier, Rüdesheimer Str. 50, 14197 Berlin. E-mail: \href{mailto:andreas.brandmaier@medicalschool-berlin.de}{\nolinkurl{andreas.brandmaier@medicalschool-berlin.de}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Psychology, MSB Medical School Berlin, Berlin, Germany\\\textsuperscript{2} Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany\\\textsuperscript{3} Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany\\\textsuperscript{4} Department of Psychology, Humboldt-Universität zu Berlin, Berlin, Germany}

\note{

A commentary on:
Gärtner, A., Leising, D., \& Schönbrodt, F. D. (2022, November 25). Responsible Research Assessment II: A specific proposal for hiring and promotion in psychology. \url{https://doi.org/10.31234/osf.io/5yexm}

Submitted to Meta-Psychology.

Participate in open peer review by commenting through hypothes.is directly on this preprint. The full editorial process of all articles under review at Meta-Psychology can be found following this link: \url{https://tinyurl.com/mp-submissions}. You will find this preprint by searching for the first author's name.

}

\abstract{%
Based on four principles of a more responsible research assessment in academic hiring and promotion processes, Gärtner, Leising, and Schönbrodt (2022) suggested an evaluation scheme for published manuscripts, reusable data sets, and research software. This commentary responds to the proposed indicators for the evaluation of research software contributions in academic hiring and promotion processes. Acknowledging the significance of research software as a critical component of modern science, we propose that an evaluation scheme must emphasize the two major dimensions of rigor and impact. Generally, we believe that research software should be recognized as valuable scientific output in academic hiring and promotion, with the hope that this incentivizes the development of more open and better research software.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Based on four principles of a more responsible research assessment in academic hiring and promotion processes (Schönbrodt et al., 2022), Gärtner et al. (2022) suggested a concrete evaluation scheme for published manuscripts, reusable data sets, and research software.
We strongly support the increased emphasis on research software as a creditable and commendable scientific contribution.
Why are research software contributions important scientific contributions?
We would like to respond with a quote from the Science Code Manifesto (Climate Code Foundation, 2011):
``Software is a cornerstone of science. Without software, twenty-first century science would be impossible''
and
``software is an essential research product, and the effort to produce, maintain, adapt, and curate code must be recognized.''
However, despite the heavy reliance on computational infrastructure, the current academic infrastructure does not adequately incentivize software development and, specifically, good software engineering practice (Baxter, Hong, Gorissen, Hetherington, \& Todorov, 2012).
In line with principle 3 of Schönbrodt et al. (2022), we suggest that criteria for research software contributions must capture two major dimensions: rigor and impact.
Impact measures whether the scholarly effort, implementation, and dissemination actually had a visible effect on the field.
Rigor means implementing high standards and best practices for ensuring transparency, correctness, and re-usability of a piece of software.
By setting a high bar of rigor in research software assessment in academic hiring and promotion, we hope to foster the creation of better software and, thus, better science.
From this perspective, we comment on some indicators of the proposed evaluation scheme for research software contributions (Table 3 of Gärtner et al., 2022).

\hypertarget{proposed-criteria}{%
\section{Proposed Criteria}\label{proposed-criteria}}

\hypertarget{id-9-citations-and-id-5-date-of-first-fully-functional-public-version}{%
\subsection{ID 9: Citations and ID 5: Date of first fully functional public version}\label{id-9-citations-and-id-5-date-of-first-fully-functional-public-version}}

Considering citations in relation to age of software seems to be inconsistent with the proposal of Schönbrodt et al. (2022), who reminded us that a core principle of the implementation of DORA is to ``abandon the use of invalid quantitative metrics of research quality and productivity in hiring and promotion'' (p.2).
It is unclear why a citation-based metric for software would be more valid than the equivalent for articles.
In fact, citations suffer from additional shortcomings when used to evaluate the impact of research software, particularly when used comparatively in the context of hiring and promotion.
For example, citing data analysis packages is much more commonly accepted than citing supporting packages (such as papaja, Aust \& Barth, 2022, used to render this article).
Further, functionality of successful, modular scientific software is ideally reused in other software packages to avoid code duplication and enable faster development of new software.
While we highly encourage reuse from both a software engineering perspective and for scientific progress, it challenges the validity of citation-based metrics for impact.
For example, consider the \texttt{NLopt} optimization suite, which counts 1,711 citations on Google scholar at the time of writing.
\texttt{NLopt} is a backbone for many scientific packages both because it implements various optimization algorithms but also because it is open-source and can inspire re-implementations of these algorithms.
One such example is the famous \texttt{lme4} package (Bates, Mächler, Bolker, \& Walker, 2014) for generalized linear mixed-effects models, which is partly based on \texttt{NLopt}.
\texttt{lme4} has more than 58,000 citations on Google scholar, yet it is unlikely that researchers will cite the underlying optimization algorithm.
For another example, the \texttt{pdc} package (Brandmaier, 2015) offers functions to cluster time series based on one specific algorithm.
The \texttt{TSclust} package (Montero \& Vilar, 2014) is a wrapper package, which imports and makes accessible functionality from the \texttt{pdc} package as well as various other clustering approaches, which is very useful from a users' perspective; however, we noticed that researchers now cite \texttt{TSclust} instead of \texttt{pdc}.

\hypertarget{id-6-date-of-most-recent-substantive-update}{%
\subsection{ID 6: Date of most recent substantive update}\label{id-6-date-of-most-recent-substantive-update}}

Both ID 5 and 6 are difficult to ascertain because it is not always clear when updates are considered `substantive' or software `fully functional'.
To assess active maintenance as an aspect of rigor, we propose to provide a simple check box, in which the author indicates whether a scientific software package is actively maintained (e.g., the software has a regular release cycle or an update within the last six months).
In addition, authors have a chance to explain why their active maintenance may be different from these guidelines.

\hypertarget{id-14-lines-of-code}{%
\subsection{ID 14: Lines of Code}\label{id-14-lines-of-code}}

We discourage the lines of code (LOC) metric to measure effort.
LOC highly depends on programming language, mastery, and personal programming style.
In particular, many LOC may simply mean that a researcher writes inefficient and repetitive code, one of the great sins of programming.
On the contrary, a feature of good software is modularity because it enables reusing functions both inside and outside the project, resulting in fewer LOC.

\hypertarget{id-7-contributor-roles-and-involvement}{%
\subsection{ID 7: Contributor Roles and Involvement}\label{id-7-contributor-roles-and-involvement}}

We support the standardized assessment of project contributor roles, similar to the Contributor Roles Taxonomy (CRediT; \url{https://credit.niso.org}).
However, we like to point out that the current evaluation schemes yields different scores for the same effort of the individual researcher depending on the size of the software project.

\hypertarget{id-8-license}{%
\subsection{ID 8: License}\label{id-8-license}}

At present, whether a piece of software is open source is not evaluated at the pre-screening stage.
However, an open license is central for assessing both rigor and (potential) impact, and should be part of the phase I assessment.
Above, we already discussed the different ways research software can have an impact --- not only by direct usage, but also by reusing software in other packages.
Open-source software makes broader impact more likely and is a prerequisite for full transparency and reproducibility (Peikert, Van Lissa, \& Brandmaier, 2021).
In addition, many aspects of rigor are impossible to evaluate for closed-source software --- for example, whether it is well-tested or bugs have been fixed.
Giving more weight to open licenses aligns well with the original article's emphasis on open science.
Therefore, we propose to penalize software if it does not adhere to an open-source license (those approved in a review process by the \href{https://opensource.org/licenses}{Open Source Initiative}), for example, by allowing them only half of the total achievable points.

\hypertarget{id-17-reusability-indicator}{%
\subsection{ID 17: Reusability Indicator}\label{id-17-reusability-indicator}}

This is one of only two criteria used in the pre-screening phase of the proposal and therefore is of central importance.
It is also important because it assesses aspects of rigor in software development: documentation, active maintenance, and testing.
However, by incorporating the size of the user base, it confounds \emph{usability} (as an aspect of rigor) with \emph{usage} (as an aspect of impact).
In addition, the criteria for the different proposed categories are not clearly defined.
For example, the difference between ``fairly extensive'' and ``extensive'' documentation is unclear.
As a result, this indicator is more of a ``gut-feeling'' indicator, roughly assessing the ``size'' of the software project.
Instead, in the following we propose to assess rigor and impact independently as primary aspects of a software contribution.

\hypertarget{assessing-rigor}{%
\section{Assessing Rigor}\label{assessing-rigor}}

We propose to use the following aspects as equally weighted indicators of rigor instead of the proposed broader re-usability indicator (item 17 in Table 3 of the proposal).

\hypertarget{tests}{%
\subsection{Tests}\label{tests}}

Tests are essential to discover incorrect functionality, investigate code scalability and reveal poor design choices.
There are a variety of useful tests, such as unit tests of sub components or tests of software functionality at a larger scale (e.g., see the \texttt{testthat} package in R, Wickham, 2011).
It is possible to quantify aspects of software testing, for example, by assessing code coverage, defined as the percentage of code lines executed during testing.
However, we believe that we should give points for software that promises that major functionality is covered by tests.
Those tests should be automated or at least open-source and reproducible.

\hypertarget{documentation}{%
\subsection{Documentation}\label{documentation}}

Just like for tests, there are different types of documentation.
For example, tutorials showcase software usage with examples, and there is application programming interface (API) documentation for individual functions and classes to enable reusing functions in other software packages.
We propose to identify relevant categories of documentation for research software, e.g., installation instructions, tutorials, API, and community guidelines (also see \url{https://joss.readthedocs.io/en/latest/review_checklist.html}), and score the presence of each of them separately.

\hypertarget{maintenance}{%
\subsection{Maintenance}\label{maintenance}}

Maintaining a software package is often more work than writing it.
This should be reflected in the assessment procedure.
We propose to score two aspects of maintenance separately, maintaining the code base (such as active bug fixing and documenting changes in logs) and maintaining the community (such as providing the possibility to report bugs, feature requests, or support requests via tickets or mailing lists).

\hypertarget{measuring-impact}{%
\section{Measuring Impact}\label{measuring-impact}}

Total citation metrics, number of users, downloads per month, GitHub stars and similar may provide a coarse measure for the impact of a software package, even though it is important to note the shortcomings we described above.

We believe that the suggested merit statement is most useful to assess impact of research software and that this should be the primary statement for committee members to evaluate if they are less concerned with the technical aspects of research software development.
In our view, researchers should be requested to indicate at least one (and up to three) research projects that directly benefited from their software contributions.
In case of recently published research software, the merit statement can be used alternatively to describe expected benefits to the field or potential use-cases of the software.
We believe this to be a fair assessment of the actual impact the specific contribution of the individual researcher had.

\hypertarget{summary}{%
\section{Summary}\label{summary}}

In sum, we are thankful to Gärtner et al. (2022) for highlighting the importance of research software contributions as scientifically valuable products.
We believe the current proposal should aim to better reflect the distinction between rigor and impact for software, similar to the guidelines proposed to evaluate journal articles.
To this end, we suggest a more fine-grained assessment of rigor, and put emphasis on the merit statement, in which software authors should argue in how far their project impacted other scientific endeavors.
Last, we hope to arrive at an evaluation scheme that incentivizes the development of scientific open-source software.
Note that the requirements of a given academic position should guide the weighting of research software when evaluating and selecting candidates.
In some cases, research software development may be central to a position and, therefore, a requirement (such as when hiring professors of statistics or research methods), but in other cases, we hope that research software development activities at least are considered as research output comparable to published papers.

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-R-papaja}{}}%
Aust, F., \& Barth, M. (2022). \emph{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}. Retrieved from \url{https://github.com/crsh/papaja}

\leavevmode\vadjust pre{\hypertarget{ref-bates2014fitting}{}}%
Bates, D., Mächler, M., Bolker, B., \& Walker, S. (2014). Fitting linear mixed-effects models using lme4. \emph{arXiv Preprint arXiv:1406.5823}.

\leavevmode\vadjust pre{\hypertarget{ref-baxter2012research}{}}%
Baxter, R., Hong, N. C., Gorissen, D., Hetherington, J., \& Todorov, I. (2012). The research software engineer. \emph{Digital Research Conference, Oxford}, 1--3.

\leavevmode\vadjust pre{\hypertarget{ref-pdc-paper}{}}%
Brandmaier, A. M. (2015). {pdc}: An {R} package for complexity-based clustering of time series. \emph{Journal of Statistical Software}, \emph{67}(5), 1--23. \url{https://doi.org/10.18637/jss.v067.i05}

\leavevmode\vadjust pre{\hypertarget{ref-manifesto}{}}%
Climate Code Foundation. (2011). \emph{Science code manifesto}. Retrieved from \url{http://web.archive.org/web/20201112032938/http://sciencecodemanifesto.org/}

\leavevmode\vadjust pre{\hypertarget{ref-gartner2022responsible}{}}%
Gärtner, A., Leising, D., \& Schönbrodt, F. (2022). \emph{Responsible research assessment {II}: A specific proposal for hiring and promotion in psychology}.

\leavevmode\vadjust pre{\hypertarget{ref-TSclust}{}}%
Montero, P., \& Vilar, J. A. (2014). {TSclust}: An {R} package for time series clustering. \emph{Journal of Statistical Software}, \emph{62}(1), 1--43. Retrieved from \url{http://www.jstatsoft.org/v62/i01/}

\leavevmode\vadjust pre{\hypertarget{ref-peikert2021reproducible}{}}%
Peikert, A., Van Lissa, C. J., \& Brandmaier, A. M. (2021). Reproducible research in r: A tutorial on how to do the same thing more than once. \emph{Psych}, \emph{3}(4), 836--867.

\leavevmode\vadjust pre{\hypertarget{ref-schonbrodt2022responsible}{}}%
Schönbrodt, F., Gärtner, A., Frank, M., Gollwitzer, M., Ihle, M., Mischkowski, D., et al.others. (2022). \emph{Responsible research assessment {I}: Implementing {DORA} for hiring and promotion in psychology}.

\leavevmode\vadjust pre{\hypertarget{ref-wickham2011}{}}%
Wickham, H. (2011). Testthat: Get started with testing. \emph{The R Journal}, \emph{3}, 5--10. Retrieved from \url{https://journal.r-project.org/archive/2011-1/RJournal_2011-1_Wickham.pdf}

\end{CSLReferences}

\end{document}

### 2023-10-04 manuscript.Rmd
add header from overleaf provided by journal

@@
 footnotelist      : no
+csquotes          : true
+babel-lang        : american
+classoption       : "jou,a4paper"
+header-includes:
+  - |
+    % Change font for whole document
+    \renewcommand{\familydefault}{\rmdefault}
+    \renewcommand*\rmdefault{bch}
 
-classoption       : "man"
-output            : papaja::apa6_pdf
+    \usepackage[english,all,portrait,draft]{draftcopy}
+    \usepackage[blocks]{authblk}
+
+    % Add watermark for preprints at the top of each page
+    % Comment out to remove preprint watermark
+    %\usepackage{draftwatermark}
+    %\SetWatermarkText{PREPRINT}
+    %\SetWatermarkScale{0.3}
+    %\SetWatermarkAngle{0}
+    %\SetWatermarkVerCenter{25pt}
+    %\SetWatermarkColor[rgb]{0.85,0.85,0.95}
+
+    % Add header with journal information
+    \usepackage{fancyhdr}
+    \renewcommand{\headrulewidth}{0pt} 	% remove vertical line
+    \setlength{\headheight}{54pt}  % increase vertical space for header
+    \fancypagestyle{firststyle}
+    {
+       \fancyhf{} % clear all header and footer fields
+
+    		\fancyhead[L]{\scriptsize  
+    		\textit{Meta-Psychology}, 2023, vol 7, MP.201X.xxx\\\href{https://doi.org/10.15626/MP.}{https://doi.org/10.15626/MP.}\\Article type: \\Published under the CC-BY4.0 license \\
+    		\smallskip
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%% Badges %%%%%%%%%%%%%%%%%%%%%
+    		\includegraphics[scale=0.2]{badges/opendata.png}
+    		\includegraphics[scale=0.2]{badges/openmaterial.png}
+    		\includegraphics[scale=0.2]{badges/preregestered.png}
+    		\includegraphics[scale=0.2]{badges/preregesteredplus.png}
+    		}
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%		
+        \fancyhead[C]{\scriptsize 
+            Open data: Yes\\
+            Open materials: Yes \\
+            Open and reproducible analysis:\\
+            Open reviews and editorial process: Yes\\
+            Preregistration: 
+            \vspace{0.5ex}
+    \\
+    		\smallskip}
+        \fancyhead[R]{\scriptsize   Edited by: \\
+            Reviewed by: \\
+            Analysis reproduced by: \\
+            All supplementary files can be accessed at OSF: \\
+            \href{https://doi.org/10.17605/OSF.IO/}{https://doi.org/10.17605/OSF.IO/}
+            \vspace{0.5ex}
+    \\
+    		\smallskip}
+    }
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    \usepackage[]{graphicx}
+    \usepackage[]{color}
+    \usepackage{tikz}  % for watermark on first page
+
+    \usepackage{alltt}
+    \usepackage[T1]{fontenc}
+    \usepackage[utf8]{inputenc}
+
+    \usepackage{amsmath}  	% for advanced math displays
+    \usepackage{booktabs}		% booktables
+
+    % With that package, tables are single spaced
+    \usepackage{setspace}
+
+    \usepackage{listings}  % for inline code blocks
+    \lstset{
+    	basicstyle=\footnotesize\ttfamily, % the size of the fonts that are used for the code
+    	breaklines=true,  % sets automatic line breaking
+    	breakatwhitespace=true % sets if automatic breaks should only happen at whitespace
+    }
+
+    \usepackage{url}	% links to headings
+    \usepackage{placeins} 	% flushes Figures before the next section starts
+    % use command \FloatBarrier
+
+    %% APA-style citations
+    \usepackage{csquotes}
+    \usepackage[backend=biber,style=apa,hyperref=true,giveninits,uniquename=init]{biblatex}
+    \DeclareLanguageMapping{american}{american-apa}
+
+    % TODO: Users must add their .bib file here
+    \addbibresource{references.bib}
+
+    % remove unwanted fields from .bib-file
+    \AtEveryBibitem{	
+      \clearfield{day}
+      \clearfield{month}
+      \clearfield{labelday}
+      \clearfield{labelmonth}
+    }
+    %\usepackage{multicol} %unneeded, apa6 already split cols
+    \raggedbottom
+
+    \usepackage{hyperref} % should be loaded last as it redefines many Latex commands
+    \hypersetup{colorlinks, urlcolor=blue, citecolor=blue}
+
+documentclass     : "apa7"
+output            :
+  papaja::apa6_pdf:
+    citation_package: biblatex
 equal_contrib: yes
-header-includes: |
-  \usepackage{hyperref}
 repro:
@@
 ```
+
+\thispagestyle{firststyle}
 

### 2023-10-04 manuscript.Rmd
remove note, add other sections

@@
       - "Writing - Review & Editing"
-
-note: |
-  A commentary on:
-  Gärtner, A., Leising, D., & Schönbrodt, F.
-D.
-(2022, November 25).
-Responsible Research Assessment II: A specific proposal for hiring and promotion in psychology.
-https://doi.org/10.31234/osf.io/5yexm
-
-  Submitted to Meta-Psychology.
-
-  Participate in open peer review by commenting through hypothes.is directly on this preprint.
-The full editorial process of all articles under review at Meta-Psychology can be found following this link: https://tinyurl.com/mp-submissions.
-You will find this preprint by searching for the first author’s name.
 
@@
 
-\newpage
+## Author Contact
 
-# References
+Put the contact information and any acknowledgements here. ORCID to all author if available. Corresponding author mandatory.
+These headings after the article but before the references are required. Feel free to add other headings (e.g., acknowledgments) if you need to.
 
-::: {#refs custom-style="Bibliography"}
-:::
+## Conflict of Interest and Funding
+
+Add a statement about conflict of interest.
+If you have no conflict of interest to declare, please state that.
+Add a statement about how the research was funded.
+If there was no specific funding please state that.
+
+## Author Contributions
+
+Explain the contribution of each authors. If you are the single author, please spell it out anyway for future proper credit of your contributions and meta-science studies. 
+The Contributor Roles Taxonomy can be helpful. https://www.casrai.org/credit.html 
+Please explain how the author names order connects to the contribution (e.g. alphabetical, ranked, or random order).
+
+## Open Science Practices
+
+ \includegraphics[scale=0.50]{badges/opendata.png}
+ \includegraphics[scale=0.50]{badges/openmaterial.png}
+ \includegraphics[scale=0.50]{badges/preregestered.png}
+ \includegraphics[scale=0.50]{badges/preregesteredplus.png}
+
+\textbf{NOTE. This will be filled out by the reproducibility team.}
+This article earned the Preregistration, Open Data and the Open Materials badge for preregistering the hypothesis and analysis before data collection, and for making the data and materials openly available.
+It has been verified that the analysis reproduced the results presented in the article.
+The entire editorial process, including the open reviews, are published in the online supplement.

### 2023-10-04 manuscript.Rmd
add School of Cognition affiliation

--- 2023-06-08 manuscript.Rmd
@@
   - id            : "4"
-    institution   : "Department of Psychology, Humboldt-Universität zu Berlin, Berlin, Germany"
+    institution   : "Max Planck School of Cognition, Leipzig, Germany"
 

### 2023-10-04 manuscript.Rmd
add aarons affiliation

@@
     institution   : "Max Planck School of Cognition, Leipzig, Germany"
+  - id            : "5"
+institution   : "Department of Imaging Neuroscience, University College London, London, UK"
 

### 2023-10-04 manuscript.Rmd
move roles from frontmatter to end

@@
     email         : "andreas.brandmaier@medicalschool-berlin.de"
-    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
-      - "Conceptualization"
-      - "Writing - Original Draft Preparation"
-      - "Writing - Review & Editing"
-      - "Supervision"
   - name          : "Maximilian Ernst"
     affiliation   : "2,4"
-    role:
-      - "Conceptualization"
-      - "Writing - Original Draft Preparation"
-      - "Writing - Review & Editing"
   - name          : "Aaron Peikert"
     affiliation   : "2,3,4"
-    role:
-      - "Conceptualization"
-      - "Writing - Original Draft Preparation"
-      - "Writing - Review & Editing"
 
@@
 
-Explain the contribution of each authors. If you are the single author, please spell it out anyway for future proper credit of your contributions and meta-science studies. 
-The Contributor Roles Taxonomy can be helpful. https://www.casrai.org/credit.html 
-Please explain how the author names order connects to the contribution (e.g. alphabetical, ranked, or random order).
+The authors made the following contributions.
+Andreas M. Brandmaier: Conceptualization, Writing - Original Draft Preparation, Writing - Review & Editing, Supervision;
+Maximilian Ernst: Conceptualization, Writing - Original Draft Preparation, Writing - Review & Editing;
+Aaron Peikert: Conceptualization, Writing - Original Draft Preparation, Writing - Review & Editing.
 

### 2023-10-04 manuscript.Rmd
add conflict of interest statement

@@
 
-Add a statement about conflict of interest.
-If you have no conflict of interest to declare, please state that.
-Add a statement about how the research was funded.
-If there was no specific funding please state that.
+We have no conflict of interest to declare.
+There was no specific funding for this research.
 

### 2023-10-04 manuscript.Rmd
make contributions a list

@@
 The authors made the following contributions.
-Andreas M. Brandmaier: Conceptualization, Writing - Original Draft Preparation, Writing - Review & Editing, Supervision;
-Maximilian Ernst: Conceptualization, Writing - Original Draft Preparation, Writing - Review & Editing;
-Aaron Peikert: Conceptualization, Writing - Original Draft Preparation, Writing - Review & Editing.
+Andreas M. Brandmaier
+: Conceptualization, Writing - Original Draft Preparation, Writing - Review & Editing, Supervision;
+Maximilian Ernst
+: Conceptualization, Writing - Original Draft Preparation, Writing - Review & Editing;
+Aaron Peikert
+: Conceptualization, Writing - Original Draft Preparation, Writing - Review & Editing.
 

### 2023-10-04 manuscript.Rmd
move correspondance

@@
     affiliation   : "1,2,3"
-    corresponding : yes    # Define only one corresponding author
+    corresponding : no    # Define only one corresponding author
     address       : "Rüdesheimer Str. 50, 14197 Berlin"
@@
 
-Put the contact information and any acknowledgements here. ORCID to all author if available. Corresponding author mandatory.
-These headings after the article but before the references are required. Feel free to add other headings (e.g., acknowledgments) if you need to.
+Correspondence concerning this article should be addressed to Andreas M. Brandmaier, Rüdesheimer Str. 50, 14197 Berlin.
+E-mail: \href{mailto:andreas.brandmaier@medicalschool-berlin.de}{\nolinkurl{andreas.brandmaier@medicalschool-berlin.de}}
 

### 2023-10-05 manuscript.Rmd
fix metadata
three final(?) changes

@@
 While we highly encourage reuse from both a software engineering perspective and for scientific progress, it challenges the validity of citation-based metrics for impact.
-For example, consider the `NLopt` optimization suite, which counts 1,711 citations on Google scholar at the time of writing.
+For example, consider the `NLopt` optimization suite, which counts 1,711 citations on Google Scholar at the time of writing.
 `NLopt` is a backbone for many scientific packages both because it implements various optimization algorithms but also because it is open-source and can inspire re-implementations of these algorithms.
-One such example is the famous `lme4` package [@bates2014fitting] for generalized linear mixed-effects models, which is partly based on `NLopt`.
+One such example is the well-known `lme4` package [@bates2014fitting] for generalized linear mixed-effects models, which is partly based on `NLopt`.
 `lme4` has more than 58,000 citations on Google scholar, yet it is unlikely that researchers will cite the underlying optimization algorithm.
@@
 To this end, we suggest a more fine-grained assessment of rigor, and put emphasis on the merit statement, in which software authors should argue in how far their project impacted other scientific endeavors.
-Last, we hope to arrive at an evaluation scheme that incentivizes the development of scientific open-source software.
+Last, we hope to arrive at an evaluation scheme that incentivizes the development of open-source research software.
 Note that the requirements of a given academic position should guide the weighting of research software when evaluating and selecting candidates.
